---
title: "Part 2: Defining Exposure, Covariates, Outcome"
author: "Maggie Li (ml4424)"
date: "8/6/2020"
output: html_document
---

```{r}
library(tidyverse)
library(tidycensus)
library(leaflet)
library(rgdal)
```

#PART 2: ASSIGN EXPOSURE/COVARIATES/OUTCOME TO COUNTIES

"Exposure" of interest = American Indian (AI) populated county (Y/N)
Outcome = PM2.5 concentration (monitor), PM2.5 concentration (model), PM2.5 coverage (monitor)

##2.1 Define AI Populated Counties using the 3 classification definitions
###2.1.1 Define Rural Cluster (Wallace et al.) Counties, extract PM2.5 values for sensitivity analysis
```{r}
# Read in county data for rural clusters & extract PM2.5 data
counties_cluster <- read.csv("Data/clustershare.csv")

# subset to just rural AI counties
native_rural_counties <- subset.data.frame(counties_cluster,
                                           cluster.name =="Rural, American Indian")

#need matching County column name with PM2.5 data 
County <- formatC(native_rural_counties$FIPS, width = 5, format = "d", flag = "0")

#add County column (updated FIPS code)
native_rural_counties <- cbind(native_rural_counties,
                               County)
native_rural_counties
#filter only to 48 states and DC
native_rural_counties <- native_rural_counties %>% 
  filter(!str_detect(County, "^02")) %>%
  filter(!str_detect(County, "^15"))
native_rural_counties
```

###2.1.2 Define Census >5% AI Counties
```{r}
#Read in and define native counties by census definition, extract PM2.5 concentrations
#Select native american counties using tidycensus: cutoff with census data of native americans > 5% of county population
census_api_key("ef9bc9f1392567620517b2f5ba86c86ebbd70d69", overwrite=TRUE, install = TRUE)
readRenviron("~/.Renviron")

#Total!!American Indian and Alaska Native alone = P003004; total population is P001001
native_pop_var <- "P003004"
total_pop <- "P001001"
native_census <- get_decennial(geography = "county", 
                               variables = native_pop_var, 
                               year = 2010)
native_census
native_census$total_pop <- get_decennial(geography = "county",
                                         variables = total_pop, 
                                         year = 2010)$value
colnames(native_census)[colnames(native_census)=="value"] <- "native_pop"
native_census$native_prop <- native_census$native_pop/native_census$total_pop
native_census

#Find counties with >5% native population; there are 203 counties.
native_census_counties <- subset.data.frame(native_census,
                                   native_prop >= 0.05)
native_census_counties <- subset(native_census_counties,
                                 select = c("GEOID", "NAME","native_prop"))
colnames(native_census_counties)[colnames(native_census_counties)=="GEOID"] <- "County"
native_census_counties <- native_census_counties %>% 
  filter(!str_detect(County, "^02")) %>%
  filter(!str_detect(County, "^15"))
native_census_counties
```

###2.1.3 Define Reservation Counties
```{r}
# Read in gdb of recognized reservations 
# Source: https://edg.epa.gov/metadata/catalog/search/resource/details.page?uuid=%7B8077CD55-74FB-4107-8047-3DEC0D55966A%7D
tribal.gdb <- "Data/Tribal_Boundaries_022117.gdb"
tribal.gdb

# List all feature classes in a file geodatabase
fc_list <- ogrListLayers(tribal.gdb)
print(fc_list)

# Read the feature class
tribal48 <- readOGR(dsn=tribal.gdb,layer="LOWER48_TRIBES")
summary(tribal48)

# Read out shapefile, to use in QGIS to find 20% overlap (commented out)
# writeOGR(obj = tribal48, 
#          dsn = "/Users/maggieli/Dropbox/Native Air Pollution Paper/Paper/CodeCheck/native-air-pollution/Data/tribal48",
#          layer = "tribal48", driver = "ESRI Shapefile")

# Do some magic in QGIS: I imported the tribal48 shapefile and the county shapefile, used the overlap function to join the layers (input layer = counties), which spits out a temporary layer showing the percent overlap of reservations in each county, for each county in the attribute table (I believe the column is called overlap_pc. I went into the properties of this layer, filtered out those with overlap_pc >= 20%, and exported this as a shapefile. I then read this in below:

# Read in cleaned counties with >= 20% area covered by reservation lands 

reservation_counties <- rgdal::readOGR("Data/tribal48/tribal20_counties.shp")
typeof(reservation_counties$GEOID)

# visualize these counties in a simple map
overlap_counties <- leaflet() %>%
addProviderTiles(providers$CartoDB.Positron) %>%
  addPolygons(data = reservation_counties, color = "#3DACF0")
overlap_counties 

# convert reservation counties spdf to dataframe
reservation_counties <- as.data.frame(reservation_counties)
reservation_counties <- reservation_counties %>% dplyr::select(GEOID) %>% dplyr::rename(County = GEOID)
reservation_counties
```

## 2.2 Define potential covariates of interest (Rural/Urban Classification, Population Density, Climate Zone, Median HH Income)
```{r}
# Rural/Urban Classification: downloaded from CDC NCHS https://www.cdc.gov/nchs/data_access/urban_rural.htm
# Key: 1 = large central metro; 2 = large fringe metro; 3 = medium metro; 4 = small metro; 5 = micropolitan (population of at least 10,000 but less than 50,000); 6 = non-core
# Basically 1-3 is urban, 4-6 is suburban/rural

urban_rural_counties_2013 <- read.csv("Data/urban_rural_counties_2013.csv")
#add leading 0; first make column vector
County <- formatC(urban_rural_counties_2013$FIPS.code, width = 5, format = "d", flag = "0")
#append new County column vector
urban_rural_counties_2013$County <- County
urban_rural_counties_2013 <- urban_rural_counties_2013 %>%
  dplyr::select(c(County, X2013.code)) %>% dplyr::rename(land_use = X2013.code)
urban_rural_counties_2013$County <- as.factor(urban_rural_counties_2013$County)
urban_rural_counties_2013$land_use <- as.factor(urban_rural_counties_2013$land_use)

# Population Density Data; downloaded from US Census 
census_pop_density <- read.csv("Data/census_2010_pop_density.csv")
head(census_pop_density)

#keep columns GCT_STUB.target.geo.id2, HD01, SUBHD0401, SUBHD0402 (population and housing density)
census_pop_density <- census_pop_density %>% subset(select = c(GCT_STUB.target.geo.id2,
                                                               HD01,
                                                               SUBHD0401, 
                                                               SUBHD0402))
#delete first three rows
census_pop_density <- census_pop_density[-c(1,2,3),]
census_pop_density <- census_pop_density %>% 
  dplyr::rename(County = GCT_STUB.target.geo.id2) %>%
  dplyr::rename(total_pop = HD01) %>%
  dplyr::rename(pop_density = SUBHD0401) %>%
  dplyr::rename(housing_density = SUBHD0402)

census_pop_density$County <- as.numeric(as.character(census_pop_density$County))
census_pop_density <- census_pop_density %>%
  filter(County <= 56045) %>% 
  filter(County > 1000)  #filter out territories
census_pop_density 
# clean county column by adding leading zero
county_cleaned <- formatC(census_pop_density$County, width = 5, format = "d", flag = "0")

census_pop_density <- cbind(census_pop_density, county_cleaned) %>%
  dplyr::select(-County)

census_pop_density$county_cleaned<- as.factor(census_pop_density$county_cleaned) 
census_pop_density <- census_pop_density%>%dplyr::rename(County = county_cleaned)

census_pop_density

# Median HH Income from 2010 ACS
hh_income <- "B19013_001"
hh_income_census <- get_acs(geography = "county",
                          variables = hh_income,
                          year = 2010)
hh_income_census <- hh_income_census %>%
  dplyr::select(c(GEOID, estimate)) %>%
  dplyr::rename(hh_income = estimate) %>%
  dplyr::rename(County = GEOID)
hh_income_census

#Climate Zone data source: https://www.ncdc.noaa.gov/monitoring-references/maps/us-climate-regions.php?hc_location=ufi

climatezonefips <- read.csv("Data/climatezonefips.csv")
climatezonefips
climatezonefips$FIPS <- as.numeric(climatezonefips$FIPS)
climatezonefips <- subset(climatezonefips,
                          select = c(Climate_Zone, FIPS))
climatezonefips <- dplyr::rename(climatezonefips,
                          State = FIPS)
climatezonefips

#EPA Regions : https://www.epa.gov/aboutepa/visiting-regional-office

eparegions <- read.csv("Data/eparegions.csv")
eparegions <- eparegions %>%
  dplyr::select(EPA_region, FIPS) %>%
  dplyr::rename(State = FIPS)
eparegions$EPA_region <- as.factor(eparegions$EPA_region)
eparegions$State <- as.numeric(eparegions$State)
eparegions

# Join altogether
covariates <- inner_join(census_pop_density,hh_income_census) %>%
  inner_join(urban_rural_counties_2013)
covariates
covariates$State <- substr(covariates$County, 0, 2)
covariates$State <- as.numeric(covariates$State)
covariates$pop_density <- as.numeric(as.character(covariates$pop_density))
covariates$housing_density <- as.numeric(as.character(covariates$housing_density))
covariates
covariates <- inner_join(covariates, eparegions)
covariates
native_census <- native_census %>% 
  dplyr::rename(County = GEOID) %>%
  dplyr::select(County, native_prop)
covariates <- covariates %>% inner_join(native_census)
covariates
```


##2.3 MAIN ANALYSIS: Define all AI/Non-AI populated counties, extract monitor/model PM2.5 concentrations & monitor coverage

After defining different AI classifications above, we collect all the counties that fit in at least one of those definitions and create a master list of predominantly AI-populated counties. We use anti-join to create a list of all non-AI populated counties.
####2.3.1 First define counties:
```{r}
# separate AI definitions from above
# native_rural_counties
# native_census_counties
# reservation_counties

# Select common county column from all three classification definitions and join into single df
census <- native_census_counties %>% dplyr::select(County)
reservation <- reservation_counties %>% dplyr::select(County)
cluster <- native_rural_counties %>% dplyr::select(County)

all_native_counties <- rbind(census, reservation, cluster)

# only keep unique values, because classification definitions overlap
all_native_counties <- unique(all_native_counties) 

# Get list of all US counties from census
all_counties <- get_decennial(geography = "county", 
                               variables = total_pop, 
                               year = 2010)
all_counties$State <- substr(all_counties$GEOID, 0, 2)
all_counties$State <- as.numeric(all_counties$State)
all_counties <- all_counties %>%
  filter(State <= 56,
         State != 2,
         State != 15) #filter out HI, AL, US territories

# County list of all non-native counties
all_counties_list <- all_counties %>% dplyr::select(GEOID) %>% 
  dplyr::rename(County = GEOID) #cleaned all_counties df to use in anti_join
all_general_counties <- anti_join(x = all_counties_list,
                                  y = all_native_counties,
                                  by= "County") %>% 
  filter(!str_detect(County, "^02"),
         !str_detect(County, "^15")) 

all_native_counties$county_type <- 1
all_general_counties$county_type <- 0

# need to filter out FIPS 46102, because it is a repeat of FIPS 46113 (Shannon County, SD was renamed Oglala Lakota County in 2015) 
all_native_counties <- all_native_counties %>% 
  filter(!County == "46102")

main_analysis_counties <- rbind(all_native_counties, all_general_counties)
main_analysis_counties
```

####2.3.2 Function to assign monitor PM2.5 values for county
```{r}
PM25_extract <- function(PM25_data, county, output_list){
  yr <- 2000
  for (year in 1:19){
    output_list[[year]] <- merge(county,
                                 PM25_data[[year]],
                                 by = "County", 
                                 all.x = TRUE) # Merge tables, Aggregate values if there are two monitors per county
    output_list[[year]]$num_mon <- ifelse(output_list[[year]]$annual_mean>=0,
                                          yes = 1, no = 0) %>% replace_na(replace = 0) #only NAs have 0 bc num_mon is binary variable (y/n)
    output_list[[year]] <- subset(output_list[[year]],
                                  select = (c(County, annual_mean, num_mon))) 
    output_list[[year]] <- aggregate(.~ County,
                                     data = output_list[[year]],
                                     FUN = sum,
                                     na.action = NULL) #sum of all monitor concentrations in county
    output_list[[year]]$annual_mean <- output_list[[year]]$annual_mean/output_list[[year]]$num_mon #divide sum by number of monitors for county avg
    output_list[[year]]$mon_present <- ifelse(output_list[[year]]$num_mon>0,
                                              yes = 1, no = 0)
    output_list[[year]]$year <- yr
    yr = yr + 1
  }
  return(output_list)
}
```

####2.3.3 Extract monitor PM2.5 concentrations and coverage:
```{r}
# Extract PM2.5 county monitor concentrations
all_native_PM25 <- list()
all_native_PM25 <- PM25_extract(PM25_all,
                                 all_native_counties,
                                 all_native_PM25) 
# Join all non-native counties with PM2.5 data
all_general_PM25 <- list()
all_general_PM25 <- PM25_extract(PM25_all,
                                 all_general_counties,
                                 all_general_PM25)
all_general_PM25
```

####2.3.4 Extract model PM2.5 concentrations:
```{r}
# extract values for main analysis AI counties
model_PM25 <- do.call(rbind, model_PM25)
model_PM25 <- model_PM25 %>% dplyr::rename(County=FIPS)
model_PM25$State <- as.numeric(model_PM25$State)
model_PM25 <- inner_join(model_PM25, covariates)
model_PM25 <- model_PM25 %>% inner_join(main_analysis_counties)

```


###2.4 SENSITIVITY ANALYSIS: Assign monitor PM2.5 concentrations for different AI classifications separately
####2.4.1 Extract monitor PM2.5 concentrations for AI=populated counties: rural cluster, census, reservation counties
```{r}
# rural cluster
native_PM25 <- list()
native_PM25 <- PM25_extract(PM25_all,
                            native_rural_counties,
                            native_PM25)

# census
native_PM25_census <- list()
native_PM25_census <- PM25_extract(PM25_all,
                         native_census_counties,
                         native_PM25_census)
native_PM25_census

# reservation
reservation_PM25 <- list()
reservation_PM25 <- PM25_extract(PM25_all,
                         reservation_counties,
                         reservation_PM25)
reservation_PM25
```

####2.4.1 Extract monitor PM2.5 concentrations for non-AI populated counties, for diff classification groups separately
#####first, define these non-AI counties
```{r}
# Define non-native counties based on reverse joining native counties with all counties
# Find all non-native census counties (by 5% threshold), all non-native cluster counties, all reservation counties; rename columns to be the same name

# standardize all dfs with single column "county"
native_census_counties_list <- native_census_counties %>% dplyr::select(County)
native_rural_counties_list <- native_rural_counties %>% dplyr::select(County)
reservation_counties <- reservation_counties #don't need to adjust, already in correct form

#get all counties that are not in these native county groups (complement)
general_counties_census <- anti_join(x = all_counties_list,
                                     y = native_census_counties,
                                     by= "County")
general_counties_cluster <- anti_join(all_counties_list,
                                     native_rural_counties_list,
                                     by = "County")
general_counties_reservations <- anti_join(x = all_counties_list,
                                           y = reservation_counties,
                                           by= "County")
```

#####now, run function to extract PM2.5 concentrations for non-native counties of all 3 classification
```{r}
#join the general counties lists with PM2.5 and PM2.5 components datasets.
#first make a list of all 3 county data frames to do a nested for loop with
general_counties <- list(general_counties_cluster,
                         general_counties_census,
                         general_counties_reservations)
#empty list of lists to store PM2.5 data across all years for different county datasets
general_PM25 <- list(cluster=list(),
                     census = list(),
                     reservations = list())

for (county in 1:length(general_counties)){
  yr <- 2000
  for (year in 1:19){
    general_PM25[[county]][[year]] <- merge(general_counties[[county]],
                               PM25_all[[year]],
                               by = "County", 
                               all.x = TRUE)
    general_PM25[[county]][[year]]$num_mon <- ifelse(general_PM25[[county]][[year]]$annual_mean>=0,
                                        yes = 1, no = 0) %>% replace_na(replace = 0)
    general_PM25[[county]][[year]] <- subset(general_PM25[[county]][[year]],
                                select = (c(County, annual_mean, num_mon)))
    general_PM25[[county]][[year]] <- aggregate(.~ County,
                                   data = general_PM25[[county]][[year]],
                                   FUN = sum,
                                   na.action = NULL)
    general_PM25[[county]][[year]]$annual_mean <- general_PM25[[county]][[year]]$annual_mean/general_PM25[[county]][[year]]$num_mon
    general_PM25[[county]][[year]]$mon_present <- ifelse(general_PM25[[county]][[year]]$num_mon>0,
                                            yes = 1, no = 0)
    general_PM25[[county]][[year]]$year <- yr
    yr = yr + 1
    }
}
length(general_PM25$reservation) #list of lists; should be 57 df total (3 classifications, 19 years each)
```

