---
title: "Part 3: Monitor Coverage & Concentrations"
author: "Maggie Li (ml4424)"
date: "10/12/2020"
output: html_document
---

## 3.1 Descriptive statistics & bar graph for monitor coverage in native/non-native counties (by different classifications)
### 3.1.1 Define fxn to calculate average # monitors per year for each county by county type, frequency plots of monitor distributions

```{r}
#1. Function to do get avg mon by county:
avg_mon <- function(data, county_type){
  monitors <- do.call(rbind, data) #create one long table with all years
  monitors$county_type <- county_type
  monitors <- subset(monitors, select = c(County, num_mon, county_type, year))
  #make long -> wide with spread, in order to have columns of each year's values, and then find the mean of these columns
  monitors <- spread(monitors, key = year, value = num_mon, fill = 0)
  monitors$avg_mon <- rowMeans(monitors[c(seq(3,21))])
  #only select the county, county_type for IDing in the bar plot, and avg # of monitor value columns
  monitors <- subset(monitors,select = c(County, county_type, avg_mon))
  return(monitors)}

#2. Frequency function to get proportion of average monitors per county across all years, using output from avg_mon:
mon_freq <- function(data, county_type){
  freq <- plyr::count(data$avg_mon) #tbl with second column that records frequencies of the diff avg_mon values
  freq$prop <- freq$freq/
    sum(freq$freq) #new proportions column
  #rename x to avg_mon
  freq <- dplyr::rename(freq, avg_mon = x)
  freq$county_type <- county_type
  return(freq)} #for native and non-native, can check this output to see if area under curve adds up to 1

#3. Function to combine frequencies of monitors for AI/non-AI counties and aggregate bins 
combine_freq <- function(native_freq, general_freq){
  all_mon <- rbind(general_freq, native_freq)
  all_mon <- all_mon %>% 
    mutate(bins = case_when(
          avg_mon == 0 ~ "0",
          avg_mon > 0 & avg_mon < 1 ~ "(0,1)", 
          avg_mon >= 1 & avg_mon < 2 ~ "[1,2)",
          avg_mon >= 2 ~ ">=2"))
  all_mon$bins <- as.factor(all_mon$bins)
  final_mon_tbl <- all_mon %>%
    group_by(bins, county_type) %>%
    dplyr::summarise(total_prop=sum(prop),
                     total_freq=sum(freq)) %>% #summarize percent monitor (prop) and total num counties (freq) values for the four bins 
    mutate(lvl_bins = factor(bins, levels = c("0","(0,1)", "[1,2)", ">=2"))) %>%
    arrange(lvl_bins, county_type)
  return(final_mon_tbl)
}  
```

### 3.1.2 MAIN ANALYSIS: Monitor coverage for all AI/non-AI counties
```{r}
native_PM25_monitors <- avg_mon(all_native_PM25, "AI-Populated County")
native_PM25_monitors
native_PM25_freq <- mon_freq(native_PM25_monitors, "AI-Populated County")
general_PM25_monitors <- avg_mon(all_general_PM25, "Non-AI Populated County")
general_PM25_freq <- mon_freq(general_PM25_monitors, "Non-AI Populated County")

all_PM25_tbl <- combine_freq(native_PM25_freq, general_PM25_freq)
# all_PM25_tbl$total_prop <- all_PM25_tbl$total_prop*100 #convert decimal-percentage to actual percentage
#check area under curve equals 1 for non-native and native
sum(general_PM25_freq$prop)
sum(native_PM25_freq$prop)

#Bar graph all counties
all_PM25_plot <- ggplot(all_PM25_tbl, 
                            aes(x=lvl_bins, y = total_prop, fill=county_type)) + 
  theme_linedraw() + #define df cols for x and y axis, select minimal theme
  geom_bar(stat="identity", position = "dodge") +
  geom_text(aes(label=total_freq), position=position_dodge(width=0.9), vjust=-0.25) + #bin sample size above bars
  scale_y_continuous(labels = function(x) paste0(x*100, "%")) + #change y-axis to percent
  theme(plot.title = element_text(size = 20),
        axis.title = element_text(size = 16),
        axis.text = element_text(size = 12),
        legend.title = element_text(size = 14),
        legend.text = element_text(size = 12)) +
  labs(title = expression(paste(PM[2.5], " Monitor Distribution, 2000-2018")),
       x = "Average number of monitors per year by county type",
       y = "Percent of Total Counties",
       fill = "County Type") 
all_PM25_plot 
all_PM25_tbl

## test changing colors
## AI = turquoise?
## Non-AI = beige
ggplot(all_PM25_tbl, aes(x=lvl_bins, y = total_prop, fill=county_type)) + 
  theme_linedraw() + #define df cols for x and y axis, select minimal theme
  geom_bar(stat="identity", position = "dodge") +
  geom_text(aes(label=total_freq), position=position_dodge(width=0.9), vjust=-0.25) + #bin sample size above bars
  scale_y_continuous(labels = function(x) paste0(x*100, "%")) + #change y-axis to percent
  theme(plot.title = element_text(size = 20),
        axis.title = element_text(size = 16),
        axis.text = element_text(size = 12),
        legend.title = element_text(size = 14),
        legend.text = element_text(size = 12)) +
  labs(title = expression(paste(PM[2.5], " Monitor Distribution, 2000-2018")),
       x = "Average number of monitors per year by county type",
       y = "Percent of Total Counties") +
  scale_fill_manual("County Type", values = c("AI-Populated County" = "#5ab4ac", "Non-AI Populated County" = "#d8b365"))
```

#### 3.1.2.1 MAIN ANALYSIS: Chi-Squared test to compare monitor distributions
```{r}
# Run just part of the above function to get table of all unique avg_mon values for AI and non-AI populated counties, and then use table function to make contingency table
native_PM25_monitors
general_PM25_monitors
all_mon <- rbind(general_PM25_freq, native_PM25_freq)
all_mon <- all_mon %>% 
  mutate(bins = case_when(
        avg_mon == 0 ~ "0",
        avg_mon > 0 & avg_mon < 1 ~ "(0,1)", 
        avg_mon >= 1 & avg_mon < 2 ~ "[1,2)",
        avg_mon >= 2 ~ ">=2"))
all_mon$bins <- as.factor(all_mon$bins)
final_mon_tbl <- all_mon %>%
  group_by(bins, county_type)
final_mon_tbl
monitor_contingency <- table(final_mon_tbl$county_type, final_mon_tbl$bins)

chisq.test(monitor_contingency)

# Conclusion: monitor distribution is not independent of AI/non-AI populated county status
```



### 3.1.3 SENSITIVITY ANALYSIS: Monitor coverage for AI/non-AI counties of the 3 different classifications
#### 3.1.3.1 Monitor coverage for census definition native/non-native counties
```{r}
#CENSUS avg monitors
native_PM25_census_monitors <- avg_mon(native_PM25_census, "AI-Populated County")
native_PM25_census_freq <- mon_freq(native_PM25_census_monitors, "AI-Populated County")
general_PM25_census_monitors <- avg_mon(general_PM25$census, "Non-AI County")
general_PM25_census_freq <- mon_freq(general_PM25_census_monitors, "Non-AI County")
census_PM25_tbl <- combine_freq(native_PM25_census_freq, general_PM25_census_freq)
census_PM25_tbl 
native_PM25_census_freq

#check area under curve equals 1 for non-native and native
sum(general_PM25_census_freq$prop)
sum(native_PM25_census_freq$prop)

#Bar graph census counties
census_PM25_plot <- ggplot(census_PM25_tbl, 
                            aes(x=lvl_bins, 
                                y = total_prop,
                                fill=county_type)) + 
  theme_linedraw() + #define df cols for x and y axis, select minimal theme
  geom_bar(stat="identity", position = "dodge") +
  geom_text(aes(label=total_freq), position=position_dodge(width=0.9), vjust=-0.25) + #bin sample size above bars
  scale_y_continuous(labels = function(x) paste0(x*100, "%")) + #change y-axis to percent
  theme(plot.title = element_text(size = 16),
        axis.title = element_text(size = 14),
        legend.title = element_text(size = 12),
        legend.text = element_text(size = 11)) +
  labs(title = expression(paste(PM[2.5], " Monitor Distribution \n in AI vs Non-AI Populated Census Counties, 2000-2018")),
       x = "Average number of monitors per year by county type",
       y = "Percent of Total Counties",
       fill = "County Type") 
census_PM25_plot 
```

#### 3.1.3.2 Monitor coverage for cluster definition native/non-native counties
```{r}
#CLUSTER avg annual mon
native_PM25_cluster_monitors <- avg_mon(native_PM25, "AI-Populated County")
native_PM25_cluster_freq <- mon_freq(native_PM25_cluster_monitors, "AI-Populated County")
general_PM25_cluster_monitors <- avg_mon(general_PM25$cluster, "Non-AI Populated County")
general_PM25_cluster_freq <- mon_freq(general_PM25_cluster_monitors, "Non-AI Populated County")

cluster_PM25_tbl <- combine_freq(general_PM25_cluster_freq, native_PM25_cluster_freq)
sum(general_PM25_cluster_freq$prop)
sum(native_PM25_cluster_freq$prop)

# need to add in "fake" extra row for >=2 natives (since 0 native counties have more than 2 monitors)
extra_row <- data.frame(">=2",
               "AI-Populated County",
               0.001,
               0,
               ">=2")
names(extra_row) <- c(names(cluster_PM25_tbl))
extra_row$county_type <- as.character(extra_row$county_type)
cluster_PM25_tbl
cluster_PM25_tbl <- bind_rows(cluster_PM25_tbl, extra_row)
cluster_PM25_tbl$lvl_bins <- as.factor(cluster_PM25_tbl$lvl_bins)
cluster_PM25_tbl
#reorder factors
cluster_PM25_tbl$lvl_bins <- factor(cluster_PM25_tbl$lvl_bins, 
                                    levels = c("0", "(0,1)", "[1,2)", ">=2"))

cluster_PM25_tbl
cluster_PM25_plot <- ggplot(cluster_PM25_tbl, 
                            aes(x=lvl_bins, 
                                y = total_prop,
                                fill=county_type)) + 
  theme_linedraw() + #define df cols for x and y axis, select minimal theme
  geom_bar(stat="identity", position = "dodge") +
  geom_text(aes(label=total_freq), position=position_dodge(width=0.9), vjust=-0.25) + #bin sample size above bars
  scale_y_continuous(labels = function(x) paste0(x*100, "%")) + #change y-axis to percent
  theme(plot.title = element_text(size = 16),
        axis.title = element_text(size = 14),
        legend.title = element_text(size = 12),
        legend.text = element_text(size = 11)) +
  labs(title = expression(paste(PM[2.5], " Monitor Distribution in Rural Clustered Counties, 2000-2018")),
       x = "Average number of monitors per year by county type",
       y = "Percent of Total Counties",
       fill = "County Type")

cluster_PM25_plot
```

####3.1.3.3 Monitor coverage for reservation definition native/non-native counties
```{r}
reservation_PM25_monitors <- avg_mon(reservation_PM25, "AI-Populated County")
reservation_PM25_freq <- mon_freq(reservation_PM25_monitors, "AI-Populated County")

general_PM25_reservations_monitors <-avg_mon(general_PM25$reservations, "General County")
general_reservation_PM25_freq <- mon_freq(general_PM25_reservations_monitors,"General County")

reservation_PM25_tbl <- combine_freq(reservation_PM25_freq,
                                     general_reservation_PM25_freq)
reservation_PM25_plot <- ggplot(reservation_PM25_tbl, 
                            aes(x=lvl_bins, 
                                y = total_prop,
                                fill=county_type)) + 
  geom_bar(stat="identity", position = "dodge") +
  labs(title = "PM2.5 Monitor Distribution in Counties containing AI Reservations \n vs. Non-AI Counties",
       x = "Average number of annual monitors per county type from 1999-2018",
       y = "Proportion of Counties",
       fill = "County Type")
sum(reservation_PM25_freq$prop)
sum(general_reservation_PM25_freq$prop)
reservation_PM25_plot
reservation_PM25_tbl
```

##3.2 Clean monitor data further to extract mean PM2.5 concentrations per county across all years, PM2.5 concentrations per county per EACH year, mean monitor count per county across all years, and monitor count per county per EACH year.

Next off, we need to clean the data further to have estimates for our Table 1, PM2.5 concentration graphs by State over time, and statistical regression models. 

###3.2.1 Compute Monitor Concentrations (averaged across all years and each annual average, per county)

####3.2.1.1 Define functions

I first define a function (drop_all_NA) that averages PM2.5 concentrations for each county across all available study years with monitors, and drops counties with NO MONITORS across all years from the AI/non-AI county dfs. The second function (combine_avg_exp) combines the two AI/non-AI county dfs into one df and gets the average PM2.5 concentration based on available monitor estimates; the native_status input is a binary indicator variable (1 = AI, 0 = non-AI) so you know which counties are which when all the counties are combined in a single df. When I run it, it's a function within a function (see below). The final output df is a table with each row entry being a county, and the average PM2.5 concentration across all study years.

The third function (exp_ctyear) iterates through all counties for each study year (no averaging) and drops county-years with NO MONITORS. The fourth function (combine_allexp), similar to the second function, combines the two AI/non-AI county dfs into one df, also joining it with the covariates df defined above, adding extra columns for state FIPS, population density & HH income quantiles, which are later used in the regression models. The final output df is a table with each row entry being a county in a given year, and the average PM2.5 concentration in that particular year; thus, counties will have multiple row entries if they had active monitor(s) in multiple years.

```{r}
#FUNCTION to extract PM2.5 exposure, averaged across all years, dropping counties with ALL NAs (i.e. counties with no monitors across all years)
drop_all_NA <- function(annual_data, native_status){
  int_data <- do.call(rbind, annual_data) #create one long table with all years
  int_data$county_type <- native_status
  int_data$county_type <- as.factor(int_data$county_type)
  int_data <- spread(int_data, key = year, value = annual_mean) #change from long to wide format (each column = year, row = county)
  int_data$avg_pm <- rowMeans(int_data[c(seq(5,23))], na.rm = TRUE) #extract mean PM2.5 concentration for each county by taking mean of rows across all years (columns)
  int_data <- subset(int_data, select = c(County, county_type, avg_pm))#only select the county, county_type for IDing in the bar plot, and avg # of monitor value columns
  int_data <- int_data %>% filter(avg_pm > 0) #drop rows (counties) with all NAs for all years (avg_pm = 0.0000)
  return(int_data)}

#FUNCTION to combine native/non-native PM2.5 annual averages
combine_avgexp <- function(native, general, fxn){
  native_exp <- fxn(native, 1)
  general_exp <- fxn(general, 0)
  full_exp <- rbind(native_exp, general_exp)
  full_exp <- full_exp %>% group_by(County,county_type) %>%
    dplyr::summarise(avg_pm_all = mean(avg_pm)) #summarizes concentration levels if multiple monitors in a county
  avgexp_full <- inner_join(full_exp, covariates)  #join with covariates data
  avgexp_full$county_type <- as.factor(avgexp_full$county_type) #convert binary indicator variable for AI/non-AI into factor
  avgexp_full$State <- substr(avgexp_full$County, 0, 2)  #add column for state based on first two characters of county FIPS code
  avgexp_full$State <- as.numeric(avgexp_full$State)  #need to recode State as numeric, for random effects
  avgexp_full <- avgexp_full %>% filter(State <= 56,
                                        State != 2,
                                        State != 15)  #filter out US territories, AL and HI
  return(avgexp_full)} 

#FUNCTION keep all exposure data, for each county year; drop any NAs
exp_ctyear <- function(annual_data, native_status){
  int_data <- do.call(rbind, annual_data) #create one long table with all years
  int_data$county_type <- native_status
  int_data$county_type <- as.factor(int_data$county_type)
  int_data <- na.omit(int_data) #remove county-years with NAs
  int_data$year <- as.factor(int_data$year)
  int_data <- subset(int_data, select = c(County, county_type, annual_mean, year))  #only select the county, county_type for AI/non-AI indicator, and avg # of monitor value columns
  return(int_data)
}

#FUNCTION to combine all county year exposure data
combine_allexp <-function(native, general, fxn){
  native_exp <- fxn(native, 1)
  general_exp <- fxn(general, 0)
  full_exp <- rbind(native_exp, general_exp)
  full_exp <- full_exp %>% group_by(County,county_type, year) %>%
    dplyr::summarise(annual_mean_all = mean(annual_mean)) #summarize exp if multiple monitors in county
  annualexp_full <- inner_join(full_exp, covariates)  #join with covariates data
  annualexp_full$county_type <- as.factor(annualexp_full$county_type)
  annualexp_full$State <- substr(annualexp_full$County, 0, 2)   #add column for state based on first two characters of county FIPS code
  annualexp_full$State <- as.numeric(annualexp_full$State)  #need to recode State as numeric, for random effects
  annualexp_full <- annualexp_full %>% filter(State <= 56,
                                        State != 2,
                                        State != 15)  #filter out territories, AL and HI
  # Create extra columns for hhinc and popd quantiles
  annualexp_full$popd_q <- cut(annualexp_full$pop_density, 
                               quantile(annualexp_full$pop_density, seq(0,1,0.1)), include.lowest = TRUE)
  annualexp_full$hhinc_q <- cut(annualexp_full$hh_income, 
                            quantile(annualexp_full$hh_income, seq(0,1,0.1)), include.lowest = TRUE)
  return(annualexp_full)
}
```

###3.2.2 Compute Monitor Coverage (average monitor count across all years per county & monitor count per county per EACH year)

####3.2.2.1 Define functions

The first function averages the number of monitors across all counties for all study years. The second function aggregates monitor count data for each county for each separate study year.

```{r}
#Join datasets to make data table with average monitors, county, native/non-native "exposure", population density
#Continuous dependent variable; Use this for t-test and linear mixed models
#Function:
avg_mon_full <- function(native, general){
  native$county_type <- 1
  general$county_type <- 0 #recode column with binary variables for model
  PM25_avgmon <- rbind(native, general)
  #join with covariates data
  avgmon_full <- inner_join(PM25_avgmon, covariates)
  avgmon_full$county_type <- as.factor(avgmon_full$county_type)
  avgmon_full
  #add column for state based on first two characters of county FIPS code
  avgmon_full$State <- substr(avgmon_full$County, 0, 2)
  #need to recode State as numeric, for random effects
  avgmon_full$State <- as.numeric(avgmon_full$State)
  avgmon_full <- avgmon_full %>% filter(State <= 56,
                                        State != 2,
                                        State != 15)
  avgmon_full <- avgmon_full %>%
    mutate(mon_present = case_when(
      avg_mon > 0 ~ 1,
      avg_mon == 0 ~ 0))
  # Create extra columns for hhinc and popd quantiles
  avgmon_full$popd_q <- cut(avgmon_full$pop_density, 
                               quantile(avgmon_full$pop_density, seq(0,1,0.1)), include.lowest = TRUE)
  avgmon_full$hhinc_q <- cut(avgmon_full$hh_income, 
                            quantile(avgmon_full$hh_income, seq(0,1,0.1)), include.lowest = TRUE)
  return(avgmon_full)}
```

```{r}
#Write function to aggregate count data across all years, also monitor present or not
count_mon <- function(native, general){
  native_count <- do.call(rbind, native)
  #binary indicator
  native_count$county_type <- 1
  general_count <- do.call(rbind, general)
  general_count$county_type <- 0
  #combine native/non-native
  count_all <- rbind(native_count,
                        general_count)
  #join with covariates data
  count_all_full <- inner_join(count_all, covariates)
  count_all_full$county_type <- as.factor(count_all_full$county_type)
  
  #add column for state based on first two characters of county FIPS code
  count_all_full$State <- substr(count_all_full$County, 0, 2)
  #need to recode State as numeric, for random effects
  count_all_full$State <- as.numeric(count_all_full$State)
  count_all_full <- count_all_full %>% filter(State <= 56,
                                        State != 2,
                                        State != 15)
  count_all_full$year <- as.factor(count_all_full$year)
  # Create extra columns for hhinc and popd quantiles
  count_all_full$popd_q <- cut(count_all_full$pop_density, 
                               quantile(count_all_full$pop_density, seq(0,1,0.1)), include.lowest = TRUE)
  count_all_full$hhinc_q <- cut(count_all_full$hh_income, 
                            quantile(count_all_full$hh_income, seq(0,1,0.1)), include.lowest = TRUE)
  return(count_all_full)}
```

####3.2.1.2 Run functions on AI/non-AI populated counties (main analysis), and separately classification definitions (sensitivity analysis)
```{r}
# MAIN ANALYSIS
  # ALL COUNTIES annual PM2.5 avgs
all_avg_exp <- combine_avgexp(all_native_PM25, all_general_PM25, drop_all_NA)
all_avg_exp
  #ALL COUNTIES PM2.5 exposures, county year 
all_ctyear_exp <- combine_allexp(all_native_PM25, all_general_PM25, exp_ctyear)

# SENSITIVITY ANALYSIS
  #CENSUS annual PM2.5 avgs
census_avg_exp <- combine_avgexp(native_PM25_census, general_PM25$census, drop_all_NA)
  #CLUSTER annual PM2.5 avgs
cluster_avg_exp<- combine_avgexp(native_PM25, general_PM25$cluster, drop_all_NA)
  #RESERVATION annual PM25 avgs 
reservation_avg_exp <- combine_avgexp(reservation_PM25, general_PM25$reservations, drop_all_NA)

  #CENSUS PM2.5 exposures county year 
census_ctyear_exp <- combine_allexp(native_PM25_census, general_PM25$census, exp_ctyear)
  #CLUSTERPM2.5 exposures county year 
cluster_ctyear_exp <- combine_allexp(native_PM25, general_PM25$cluster, exp_ctyear)
  #RESERVATION PM2.5 exposures county year
reservation_ctyear_exp <- combine_allexp(reservation_PM25, general_PM25$reservations, exp_ctyear)

# DFs of MEAN annual PM2.5 concentrations across all study years for each county with data
all_avg_exp
census_avg_exp
cluster_avg_exp
reservation_avg_exp

# DFs of PM2.5 concentrations for each COUNTY-YEAR
all_ctyear_exp
census_ctyear_exp
cluster_ctyear_exp
reservation_ctyear_exp
```

####3.2.1.2 Run functions on AI/non-AI populated counties (main analysis), and separately classification definitions (sensitivity analysis)
```{r}
# MAIN ANALYSIS
  #Average monitor count data
all_avgmon_full <- avg_mon_full(native_PM25_monitors,
                                general_PM25_monitors)

  #Monitor count data for each county-year
all_native_count_full <- count_mon(all_native_PM25, all_general_PM25)

all_avgmon_full
all_native_count_full

# SENSITIVITY ANALYSIS
  #1) Census Average Monitor Count
census_avgmon_full <- avg_mon_full(native_PM25_census_monitors, 
                              general_PM25_census_monitors)
  # Census Monitor Count for each county-year
census_count_full <- count_mon(native_PM25_census, general_PM25$census)

  #2) Cluster Average Monitor Count
cluster_avgmon_full <- avg_mon_full(native_PM25_cluster_monitors, 
                               general_PM25_cluster_monitors)
  #Cluster Monitor Count for each county-year
cluster_count_full <- count_mon(native_PM25, general_PM25$cluster)

  #3) Reservation Average Monitor Count
reservation_avgmon_full <- avg_mon_full(reservation_PM25_monitors,
                            general_PM25_reservations_monitors)
  #Reservation Monitor Count for each county-year
reservation_count_full <- count_mon(reservation_PM25, general_PM25$reservations)

census_avgmon_full
census_count_full

cluster_avgmon_full
cluster_count_full

reservation_avgmon_full
reservation_count_full
```
