---
title: 'Part 4: Table 1'
author: "Maggie Li (ml4424)"
date: "10/12/2020"
output: html_document
---

```{r}
library(tidyverse)
library(xtable)
require(data.table) #need to convert for latex export
```

## Table 1: Descriptive Statistics by County for all counties, all counties with monitors, rural counties, rural counties with monitors (AI/Non-AI)

Characteristics include: number of counties, mean annual PM2.5 monitor count, mean monitor PM2.5 concentration, mean model PM2.5 concentration, mean population density, mean household income.

I created two tables, one with all counties and counties with monitors (first chunk), the next with rural counties and rural counties with monitors (second chunk).

August 4 update: fixed the mean annual concentrations by monitor and SD

### All Counties 
```{r}
# First table has columns for all counties and all counties with monitors
# Join 
all_native_counties_full <- inner_join(all_native_counties, covariates)
all_general_counties_full <- inner_join(all_general_counties, covariates)
#Number of counties
num_counties_row_main <- data.frame(nrow(all_native_counties_full),
                                    nrow(all_general_counties_full),
                                    nrow(all_avg_exp %>% filter(county_type==1)),
                                    nrow(all_avg_exp %>% filter(county_type==0)))
names(num_counties_row_main) <- c("AI-Populated County", 
                                  "Non-AI-Populated County",
                                  "AI-Populated County with monitors",
                                  "Non-AI-Populated County with monitors")

#Mean monitor count and SD
native_avgmon_full <- all_avgmon_full %>% filter(county_type == 1) 
general_avgmon_full <- all_avgmon_full %>% filter(county_type == 0)
native_monitors_avgmon_full <- all_avgmon_full %>% filter(county_type == 1,
                                                           avg_mon > 0) 
general_monitors_avgmon_full <- all_avgmon_full %>% filter(county_type == 0,
                                                           avg_mon >0)
mean_mon_row_main <- data.frame(mean(native_avgmon_full$avg_mon),
                                mean(general_avgmon_full$avg_mon),
                                mean(native_monitors_avgmon_full$avg_mon),
                                mean(general_monitors_avgmon_full$avg_mon))
names(mean_mon_row_main) <- names(num_counties_row_main)

sd_mon_row_main <- data.frame(sd(native_avgmon_full$avg_mon),
                                sd(general_avgmon_full$avg_mon),
                                sd(native_monitors_avgmon_full$avg_mon),
                                sd(general_monitors_avgmon_full$avg_mon))
names(sd_mon_row_main) <- names(num_counties_row_main)

#Mean Annual Concentrations by Monitors & SD
native_avg_exp <- all_avg_exp %>% filter(county_type == 1)
general_avg_exp <- all_avg_exp %>% filter(county_type == 0)

avg_PM25_row_main <-data.frame(mean(native_avg_exp$avg_pm_all),
                               mean(general_avg_exp$avg_pm_all),
                               mean(native_avg_exp$avg_pm_all),
                               mean(general_avg_exp$avg_pm_all))
names(avg_PM25_row_main) <- names(num_counties_row_main) 
sd_PM25_row_main <- data.frame(sd(native_avg_exp$avg_pm_all),
                               sd(general_avg_exp$avg_pm_all),
                               sd(native_avg_exp$avg_pm_all),
                               sd(general_avg_exp$avg_pm_all))
names(sd_PM25_row_main) <- names(num_counties_row_main) 

#Mean Annual Concentrations by Model & SD
  # First, obtain average model data
avg_modelPM25 <- spread(model_PM25, key = Year, value = PM25)


avg_modelPM25$avg_modelpm <- rowMeans(avg_modelPM25[c(seq(11,29))], na.rm = TRUE) #get mean of all study years 2000-2018
avg_modelPM25 

native_modelexp <- avg_modelPM25 %>% filter(county_type == 1)
general_modelexp <- avg_modelPM25 %>% filter(county_type == 0)

avg_modelPM25$county_type <- as.factor(avg_modelPM25$county_type) #need to convert into a factor to join with the other dataset

# modeled pm2.5 for counties with monitors
native_monitor_modelexp <- avg_modelPM25 %>% inner_join(native_avg_exp) %>%
  filter(county_type == 1)

general_monitor_modelexp <- avg_modelPM25 %>% inner_join(general_avg_exp) %>%
  filter(county_type == 0)
general_monitor_modelexp

# make avg model row
avg_modelPM25_row_main <-data.frame(mean(native_modelexp$avg_modelpm),
                               mean(general_modelexp$avg_modelpm),
                               mean(native_monitor_modelexp$avg_modelpm),
                               mean(general_monitor_modelexp$avg_modelpm))
names(avg_modelPM25_row_main) <- names(num_counties_row_main) 

sd_modelPM25_row_main <- data.frame(sd(native_modelexp$avg_modelpm),
                               sd(general_modelexp$avg_modelpm),
                               sd(native_monitor_modelexp$avg_modelpm),
                               sd(general_monitor_modelexp$avg_modelpm))
names(sd_modelPM25_row_main) <- names(num_counties_row_main) 

#covariate mean and SD rows
  # population density
popdensity_row_main <- data.frame(mean(native_avgmon_full$pop_density),
                                  mean(general_avgmon_full$pop_density),
                                  mean(native_monitors_avgmon_full$pop_density),
                                  mean(general_monitors_avgmon_full$pop_density))
names(popdensity_row_main) <- names(num_counties_row_main) 

sd_popden_main <- data.frame(sd(native_avgmon_full$pop_density),
                                  sd(general_avgmon_full$pop_density),
                                  sd(native_monitors_avgmon_full$pop_density),
                                  sd(general_monitors_avgmon_full$pop_density))
names(sd_popden_main) <- names(num_counties_row_main) 

  # hh income
hhinc_row_main <- data.frame(mean(native_avgmon_full$hh_income),
                             mean(general_avgmon_full$hh_income),
                             mean(native_monitors_avgmon_full$hh_income),
                             mean(general_monitors_avgmon_full$hh_income))
names(hhinc_row_main) <- names(num_counties_row_main) 
sd_hhinc_main <- data.frame(sd(native_avgmon_full$hh_income),
                             sd(general_avgmon_full$hh_income),
                             sd(native_monitors_avgmon_full$hh_income),
                             sd(general_monitors_avgmon_full$hh_income))
names(sd_hhinc_main) <- names(num_counties_row_main) 


#round mean rows
num_counties_row_main[] <- lapply(num_counties_row_main,as.integer)
mean_mon_row_main <- format(round(mean_mon_row_main, 1), nsmall = 1)
avg_PM25_row_main <- format(round(avg_PM25_row_main, 1), nsmall = 1)
avg_modelPM25_row_main <-format(round(avg_modelPM25_row_main, 1), nsmall = 1)
popdensity_row_main[] <- lapply(popdensity_row_main,as.integer)
hhinc_row_main[] <- lapply(hhinc_row_main,as.integer)

#round sd rows
sd_mon_row_main <- format(round(sd_mon_row_main, 1), nsmall = 1)
sd_PM25_row_main <- format(round(sd_PM25_row_main, 1), nsmall = 1)
sd_modelPM25_row_main <-format(round(sd_modelPM25_row_main, 1), nsmall = 1)
sd_popden_main[] <- lapply(sd_popden_main, as.integer)
sd_hhinc_main[] <- lapply(sd_hhinc_main, as.integer)


sd_mon_row_main
sd_PM25_row_main
sd_popden_main
sd_hhinc_main

#make list of rows
rows_list_main <- list(num_counties_row_main,
      mean_mon_row_main,
      avg_PM25_row_main,
      avg_modelPM25_row_main,
      popdensity_row_main,
      hhinc_row_main)
rows_list_main
#rbind all rows together
table_1_main <- do.call(rbind, rows_list_main)
table_1_main
row.names(table_1_main) <- c("Number of Counties",
                             "Average Monitor Count per year",
                        "Mean Monitor PM2.5 Concentration",
                        "Mean Model PM2.5 Concentrations",
                        "Population Density",
                        "HH Income")

test <- setDT(table_1_main)
main_table1 <- xtable(test)
main_table1

#sd table (to manually input to table 1)
row_list_sd <- list(sd_mon_row_main,
                    sd_PM25_row_main,
                    sd_modelPM25_row_main,
                    sd_popden_main,
                    sd_hhinc_main)
table_1_sd <- do.call(rbind,row_list_sd)
row.names(table_1_sd) <- c("SD Mean Monitor Count",
                           "SD Monitor PM2.5",
                        "SD Model PM2.5",
                        "Population Density",
                        "HH Income")
test_sd <- setDT(table_1_sd)
main_table_sd <- xtable(test_sd)
main_table_sd
```

### Rural Counties
Note: I use the same variable names as above in intermediate steps for calculating each of the characteristics (overwriting values), so please run each chunk in their entirety to ensure the output tables

**Definition of Rural Counties:** Micropolitan and Non-Core
Source: https://www.hrsa.gov/rural-health/about-us/definition/index.html#:~:text=All%20counties%20that%20are%20not,as%20either%20Metro%20or%20Micro

"Micropolitan counties are considered non-Metropolitan or rural along with all counties that are not classified as either Metro or Micro. "

```{r}
# Table 1.2: Second table has columns for rural counties and rural counties with monitors
rural_native <- all_native_counties_full %>% filter(as.numeric(land_use) >= 5)
rural_general <- all_general_counties_full %>% filter(as.numeric(land_use) >= 5)

#Number of counties
num_counties_row_main <- data.frame(nrow(rural_native),
                                    nrow(rural_general),
                                    nrow(all_avg_exp %>% filter(as.numeric(land_use) >= 5,
                                                                county_type==1)),
                                    nrow(all_avg_exp %>% filter(as.numeric(land_use) >= 5,
                                                                county_type==0)))
names(num_counties_row_main) <- c("Rural AI-Populated County", 
                                  "Rural Non-AI-Populated County",
                                  "Rural AI-Populated County with monitors",
                                  "Rural Non-AI-Populated County with monitors")

# Rural counties: Mean monitor count & SD
rural_native_avgmon <- native_avgmon_full %>% filter(as.numeric(land_use) >= 5)
rural_general_avgmon <- general_avgmon_full %>% filter(as.numeric(land_use) >= 5)
rural_native_monitors_avgmon <- rural_native_avgmon %>% filter(county_type == 1,
                                                           avg_mon > 0) 
rural_general_monitors_avgmon <- rural_general_avgmon %>% filter(county_type == 0,
                                                           avg_mon >0)
mean_mon_row_main <- data.frame(mean(rural_native_avgmon$avg_mon),
                                mean(rural_general_avgmon$avg_mon),
                                mean(rural_native_monitors_avgmon$avg_mon),
                                mean(rural_general_monitors_avgmon$avg_mon))
names(mean_mon_row_main) <- names(num_counties_row_main)

sd_mon_row_main <- data.frame(sd(rural_native_avgmon$avg_mon),
                                sd(rural_general_avgmon$avg_mon),
                                sd(rural_native_monitors_avgmon$avg_mon),
                                sd(rural_general_monitors_avgmon$avg_mon))
names(sd_mon_row_main) <- names(num_counties_row_main)

# Rural counties: monitor mean concentrations & SD
rural_native_avgexp <- native_avg_exp %>% filter(as.numeric(land_use) >= 5)
rural_general_avgexp <- general_avg_exp %>% filter(as.numeric(land_use) >= 5)

# new variables for PM2.5 concentrations in counties with monitors
monitors_rural_avgexp <- rural_native_avgexp
monitors_general_avgexp <- rural_general_avgexp

avg_PM25_row_main <-data.frame(mean(rural_native_avgexp$avg_pm_all),
                               mean(rural_general_avgexp$avg_pm_all),
                               mean(monitors_rural_avgexp$avg_pm_all),
                               mean(monitors_general_avgexp$avg_pm_all))
names(avg_PM25_row_main) <- names(num_counties_row_main) 

sd_PM25_row_main <- data.frame(sd(rural_native_avgexp$avg_pm_all),
                               sd(rural_general_avgexp$avg_pm_all),
                               sd(monitors_rural_avgexp$avg_pm_all),
                               sd(monitors_general_avgexp$avg_pm_all))
names(sd_PM25_row_main) <- names(num_counties_row_main) 
# Rural counties: model mean concentrations & SD
rural_native_modelexp <- avg_modelPM25 %>% filter(county_type == 1,
                                                  as.numeric(land_use) >= 5)
rural_general_modelexp <- avg_modelPM25 %>% filter(county_type == 0,
                                                   as.numeric(land_use) >= 5)
rural_native_monitor_modelexp <- avg_modelPM25 %>% inner_join(native_avg_exp) %>%
  filter(county_type == 1, as.numeric(land_use) >= 5)
rural_general_monitor_modelexp <- avg_modelPM25 %>% inner_join(monitors_general_avgexp) %>%
  filter(county_type == 0, as.numeric(land_use) >= 5)


avg_modelPM25_row_main <-data.frame(mean(rural_native_modelexp$avg_modelpm),
                               mean(rural_general_modelexp$avg_modelpm),
                               mean(rural_native_monitor_modelexp$avg_modelpm),
                               mean(rural_general_monitor_modelexp$avg_modelpm))
names(avg_modelPM25_row_main) <- names(num_counties_row_main) 

sd_modelPM25_row_main <- data.frame(sd(rural_native_modelexp$avg_modelpm),
                               sd(rural_general_modelexp$avg_modelpm),
                               sd(rural_native_monitor_modelexp$avg_modelpm),
                               sd(rural_general_monitor_modelexp$avg_modelpm))
names(sd_modelPM25_row_main) <- names(num_counties_row_main) 

#covariate mean rows! &SD
  # population density
popdensity_row_main <- data.frame(mean(rural_native_avgmon$pop_density),
                                  mean(rural_general_avgmon$pop_density),
                                  mean(rural_native_monitors_avgmon$pop_density),
                                  mean(rural_general_monitors_avgmon$pop_density))
names(popdensity_row_main) <- names(num_counties_row_main) 

sd_popden_main <- data.frame(sd(rural_native_avgmon$pop_density),
                                  sd(rural_general_avgmon$pop_density),
                                  sd(rural_native_monitors_avgmon$pop_density),
                                  sd(rural_general_monitors_avgmon$pop_density))
names(sd_popden_main) <- names(num_counties_row_main) 

  # hh income
hhinc_row_main <- data.frame(mean(rural_native_avgmon$hh_income),
                             mean(rural_general_avgmon$hh_income),
                             mean(rural_native_monitors_avgmon$hh_income),
                             mean(rural_general_monitors_avgmon$hh_income))
names(hhinc_row_main) <- names(num_counties_row_main) 
sd_hhinc_main <- data.frame(sd(rural_native_avgmon$hh_income),
                             sd(rural_general_avgmon$hh_income),
                             sd(rural_native_monitors_avgmon$hh_income),
                             sd(rural_general_monitors_avgmon$hh_income))
names(sd_hhinc_main) <- names(num_counties_row_main) 


#round mean rows first
num_counties_row_main[] <- lapply(num_counties_row_main,as.integer)
mean_mon_row_main <- format(round(mean_mon_row_main, 1), nsmall = 1)
avg_PM25_row_main <- format(round(avg_PM25_row_main, 1), nsmall = 1)
avg_modelPM25_row_main <-format(round(avg_modelPM25_row_main, 1), nsmall = 1)
popdensity_row_main[] <- lapply(popdensity_row_main,as.integer)
hhinc_row_main[] <- lapply(hhinc_row_main,as.integer)

#round sd rows
sd_mon_row_main <- format(round(sd_mon_row_main, 1), nsmall = 1)
sd_PM25_row_main <- format(round(sd_PM25_row_main, 1), nsmall = 1)
sd_modelPM25_row_main <-format(round(sd_modelPM25_row_main, 1), nsmall = 1)
sd_popden_main[] <- lapply(sd_popden_main, as.integer)
sd_hhinc_main[] <- lapply(sd_hhinc_main, as.integer)


sd_mon_row_main
sd_PM25_row_main
sd_popden_main
sd_hhinc_main

#make list of rows
rows_list_main_rural <- list(num_counties_row_main,
      mean_mon_row_main,
      avg_PM25_row_main,
      avg_modelPM25_row_main,
      popdensity_row_main,
      hhinc_row_main)
rows_list_main_rural
#rbind all rows together
table_1_main_rural <- do.call(rbind, rows_list_main_rural)
table_1_main_rural
row.names(table_1_main_rural) <- c("Number of Counties",
                             "Average Monitor Count per year",
                        "Mean Monitor PM2.5 Concentration",
                        "Mean Model PM2.5 Concentrations",
                        "Population Density",
                        "HH Income")
table1_rural_dt <- setDT(table_1_main_rural)
main_table1_rural <- xtable(table1_rural_dt)
main_table1_rural

#sd table (to manually input to table 1)
row_list_sd_rural <- list(sd_mon_row_main,
                    sd_PM25_row_main,
                    sd_modelPM25_row_main,
                    sd_popden_main,
                    sd_hhinc_main)
table_1_sd_rural <- do.call(rbind,row_list_sd_rural)
row.names(table_1_sd_rural) <- c("SD Mean Monitor Count",
                           "SD Monitor PM2.5",
                        "SD Model PM2.5",
                        "Population Density",
                        "HH Income")
table1_sd_rural <- setDT(table_1_sd_rural)
main_table_sd_rural <- xtable(table1_sd_rural)
main_table_sd_rural
```

I manually joined the two (all counties and rural counties) and just messed with the formatting in LaTeX a bit to make it all into one table. :)