---
title: 'Part 4: Table 1'
author: "Maggie Li (ml4424)"
date: "10/12/2020"
output: html_document
---

```{r}
library(tidyverse)
library(xtable)
require(data.table) #need to convert for latex export
```


## Table 1: Descriptive Statistics by County for all counties, all counties with monitors, rural counties, rural counties with monitors (AI/Non-AI)

Characteristics include: number of counties, % AI/AN, measured PM2.5 concentration, model PM2.5 concentration, population density, median household income, ADI

Table 1 includes statistics for two sets of counties: (1) all study counties and study counties with monitors (first chunk), (2) all rural study counties and rural study counties with monitors (second chunk). Each code chunk create one table of mean values and one table of SD values for the different characteristics.


```{r ADI}
adi_blckgrp = read_delim("Data/adi-download/US_2015_ADI_Census Block Group_v3.1.txt",
                         delim = ",")
substr(adi_blckgrp$FIPS, 1, 5)

adi_county = adi_blckgrp %>% 
  mutate(County = substr(FIPS, 1, 5),
         ADI_NATRANK = as.numeric(ADI_NATRANK), 
         ADI_STATERNK = as.numeric(ADI_STATERNK)) %>% 
  dplyr::select(County, ADI_NATRANK, ADI_STATERNK) %>% 
  group_by(County) %>% 
  summarize(mean_nat_adi = mean(ADI_NATRANK, na.rm = TRUE),
            mean_state_adi = mean(ADI_STATERNK, na.rm = TRUE))
```


### All Counties 
```{r}
# First table has columns for all counties and all counties with monitors
# Join 
all_native_counties_full <- inner_join(all_native_counties, covariates) %>% 
  inner_join(adi_county)
all_general_counties_full <- inner_join(all_general_counties, covariates) %>% 
  filter(!str_detect(County, "51515")) %>% 
  inner_join(adi_county)

all_avgmon_full = all_avgmon_full %>% 
  inner_join(adi_county %>% 
               mutate(County = str_replace(County, "46102", "46113"))) %>% 
  filter(!str_detect(County, "51515")) # drop to have 3108 counties

all_native_counties_full
all_general_counties_full
#Number of counties
num_counties_row_main <- data.frame(nrow(all_native_counties_full),
                                    nrow(all_general_counties_full),
                                    nrow(all_avg_exp %>% filter(county_type==1)),
                                    nrow(all_avg_exp %>% filter(county_type==0)))
names(num_counties_row_main) <- c("AI-Populated County", 
                                  "Non-AI-Populated County",
                                  "AI-Populated County with monitors",
                                  "Non-AI-Populated County with monitors")

#Mean monitor count and SD
native_avgmon_full <- all_avgmon_full %>% filter(county_type == 1) 
general_avgmon_full <- all_avgmon_full %>% filter(county_type == 0)
native_monitors_avgmon_full <- all_avgmon_full %>% filter(county_type == 1,
                                                           avg_mon > 0) 
general_monitors_avgmon_full <- all_avgmon_full %>% filter(county_type == 0,
                                                           avg_mon >0)
mean_mon_row_main <- data.frame(mean(native_avgmon_full$avg_mon),
                                mean(general_avgmon_full$avg_mon),
                                mean(native_monitors_avgmon_full$avg_mon),
                                mean(general_monitors_avgmon_full$avg_mon))
names(mean_mon_row_main) <- names(num_counties_row_main)

sd_mon_row_main <- data.frame(sd(native_avgmon_full$avg_mon),
                                sd(general_avgmon_full$avg_mon),
                                sd(native_monitors_avgmon_full$avg_mon),
                                sd(general_monitors_avgmon_full$avg_mon))
names(sd_mon_row_main) <- names(num_counties_row_main)

#Mean Annual Concentrations by Monitors & SD
native_avg_exp <- all_avg_exp %>% filter(county_type == 1)
general_avg_exp <- all_avg_exp %>% filter(county_type == 0)

avg_PM25_row_main <-data.frame(mean(native_avg_exp$avg_pm_all),
                               mean(general_avg_exp$avg_pm_all),
                               mean(native_avg_exp$avg_pm_all),
                               mean(general_avg_exp$avg_pm_all))
names(avg_PM25_row_main) <- names(num_counties_row_main) 
sd_PM25_row_main <- data.frame(sd(native_avg_exp$avg_pm_all),
                               sd(general_avg_exp$avg_pm_all),
                               sd(native_avg_exp$avg_pm_all),
                               sd(general_avg_exp$avg_pm_all))
names(sd_PM25_row_main) <- names(num_counties_row_main) 

#Mean Annual Concentrations by Model & SD
  # First, obtain average model data
avg_modelPM25 <- spread(model_PM25, key = Year, value = PM25)


avg_modelPM25$avg_modelpm <- rowMeans(avg_modelPM25[c(seq(11,29))], na.rm = TRUE) #get mean of all study years 2000-2018
avg_modelPM25 

native_modelexp <- avg_modelPM25 %>% filter(county_type == 1)
general_modelexp <- avg_modelPM25 %>% filter(county_type == 0)

avg_modelPM25$county_type <- as.factor(avg_modelPM25$county_type) #need to convert into a factor to join with the other dataset

# modeled pm2.5 for counties with monitors
native_monitor_modelexp <- avg_modelPM25 %>% inner_join(native_avg_exp) %>%
  filter(county_type == 1)

general_monitor_modelexp <- avg_modelPM25 %>% inner_join(general_avg_exp) %>%
  filter(county_type == 0)
general_monitor_modelexp

# make avg model row
avg_modelPM25_row_main <-data.frame(mean(native_modelexp$avg_modelpm),
                               mean(general_modelexp$avg_modelpm),
                               mean(native_monitor_modelexp$avg_modelpm),
                               mean(general_monitor_modelexp$avg_modelpm))
names(avg_modelPM25_row_main) <- names(num_counties_row_main) 

sd_modelPM25_row_main <- data.frame(sd(native_modelexp$avg_modelpm),
                               sd(general_modelexp$avg_modelpm),
                               sd(native_monitor_modelexp$avg_modelpm),
                               sd(general_monitor_modelexp$avg_modelpm))
names(sd_modelPM25_row_main) <- names(num_counties_row_main) 

#covariate mean and SD rows

  # population density
popdensity_row_main <- data.frame(mean(native_avgmon_full$pop_density),
                                  mean(general_avgmon_full$pop_density),
                                  mean(native_monitors_avgmon_full$pop_density),
                                  mean(general_monitors_avgmon_full$pop_density))
names(popdensity_row_main) <- names(num_counties_row_main) 

sd_popden_main <- data.frame(sd(native_avgmon_full$pop_density),
                                  sd(general_avgmon_full$pop_density),
                                  sd(native_monitors_avgmon_full$pop_density),
                                  sd(general_monitors_avgmon_full$pop_density))
names(sd_popden_main) <- names(num_counties_row_main) 

  # hh income
hhinc_row_main <- data.frame(mean(native_avgmon_full$hh_income),
                             mean(general_avgmon_full$hh_income),
                             mean(native_monitors_avgmon_full$hh_income),
                             mean(general_monitors_avgmon_full$hh_income))
names(hhinc_row_main) <- names(num_counties_row_main) 
sd_hhinc_main <- data.frame(sd(native_avgmon_full$hh_income),
                             sd(general_avgmon_full$hh_income),
                             sd(native_monitors_avgmon_full$hh_income),
                             sd(general_monitors_avgmon_full$hh_income))
names(sd_hhinc_main) <- names(num_counties_row_main) 


  # ADI
adi_row_main_25 <- data.frame(quantile(native_avgmon_full$mean_nat_adi, 0.25, na.rm = T),
                                  quantile(general_avgmon_full$mean_nat_adi, 0.25, na.rm = T),
                                  quantile(native_monitors_avgmon_full$mean_nat_adi, 0.25, na.rm = T),
                                  quantile(general_monitors_avgmon_full$mean_nat_adi, 0.25, na.rm = T))
names(adi_row_main_25) <- names(num_counties_row_main) 

adi_main_75 <- data.frame(quantile(native_avgmon_full$mean_nat_adi, 0.75, na.rm = T),
                                  quantile(general_avgmon_full$mean_nat_adi, 0.75, na.rm = T),
                                  quantile(native_monitors_avgmon_full$mean_nat_adi, 0.75, na.rm = T),
                                  quantile(general_monitors_avgmon_full$mean_nat_adi, 0.75, na.rm = T))
names(adi_main_75) <- names(num_counties_row_main) 

#round mean rows
num_counties_row_main[] <- lapply(num_counties_row_main,as.integer)
mean_mon_row_main <- format(round(mean_mon_row_main, 1), nsmall = 1)
avg_PM25_row_main <- format(round(avg_PM25_row_main, 1), nsmall = 1)
avg_modelPM25_row_main <-format(round(avg_modelPM25_row_main, 1), nsmall = 1)
popdensity_row_main[] <- lapply(popdensity_row_main,as.integer)
hhinc_row_main[] <- lapply(hhinc_row_main,as.integer)
adi_row_main_25[] <- format(round(adi_row_main_25, 1), nsmall = 1)

#round sd rows
sd_mon_row_main <- format(round(sd_mon_row_main, 1), nsmall = 1)
sd_PM25_row_main <- format(round(sd_PM25_row_main, 1), nsmall = 1)
sd_modelPM25_row_main <-format(round(sd_modelPM25_row_main, 1), nsmall = 1)
sd_popden_main[] <- lapply(sd_popden_main, as.integer)
sd_hhinc_main[] <- lapply(sd_hhinc_main, as.integer)
adi_main_75[] <- format(round(adi_main_75, 1), nsmall = 1)

#make list of rows
rows_list_main <- list(num_counties_row_main,
      mean_mon_row_main,
      avg_PM25_row_main,
      avg_modelPM25_row_main,
      popdensity_row_main,
      hhinc_row_main,
      adi_row_main_25)
rows_list_main
#rbind all rows together
table_1_main <- do.call(rbind, rows_list_main)
table_1_main
row.names(table_1_main) <- c("Number of Counties",
                             "Average Monitor Count per year",
                        "Mean Monitor PM2.5 Concentration",
                        "Mean Model PM2.5 Concentrations",
                        "Population Density",
                        "HH Income",
                        "ADI 25th Percentile")
table_1_main
# test <- setDT(table_1_main)
# main_table1 <- xtable(test)
# main_table1

#sd table (to manually input to table 1)
row_list_sd <- list(sd_mon_row_main,
                    sd_PM25_row_main,
                    sd_modelPM25_row_main,
                    sd_popden_main,
                    sd_hhinc_main,
                    adi_main_75)
table_1_sd <- do.call(rbind,row_list_sd)
row.names(table_1_sd) <- c("SD Mean Monitor Count",
                           "SD Monitor PM2.5",
                        "SD Model PM2.5",
                        "Population Density",
                        "HH Income",
                        "ADI 75th Percentile")
table_1_sd
# test_sd <- setDT(table_1_sd)
# main_table_sd <- xtable(test_sd)
# main_table_sd
```


### Rural Counties
Note: I use the same variable names as above in intermediate steps for calculating each of the characteristics (overwriting values), so please run each chunk in their entirety to ensure the output tables

**Definition of Rural Counties:** Micropolitan and Non-Core
Source: https://www.hrsa.gov/rural-health/about-us/definition/index.html#:~:text=All%20counties%20that%20are%20not,as%20either%20Metro%20or%20Micro

"Micropolitan counties are considered non-Metropolitan or rural along with all counties that are not classified as either Metro or Micro. "

```{r}
# Table 1.2: Second table has columns for rural counties and rural counties with monitors
rural_native <- all_native_counties_full %>% filter(as.numeric(land_use) >= 5)
rural_general <- all_general_counties_full %>% filter(as.numeric(land_use) >= 5)

#Number of counties
num_counties_row_main <- data.frame(nrow(rural_native),
                                    nrow(rural_general),
                                    nrow(all_avg_exp %>% filter(as.numeric(land_use) >= 5,
                                                                county_type==1)),
                                    nrow(all_avg_exp %>% filter(as.numeric(land_use) >= 5,
                                                                county_type==0)))
names(num_counties_row_main) <- c("Rural AI-Populated County", 
                                  "Rural Non-AI-Populated County",
                                  "Rural AI-Populated County with monitors",
                                  "Rural Non-AI-Populated County with monitors")

# Rural counties: Mean monitor count & SD
rural_native_avgmon <- native_avgmon_full %>% filter(as.numeric(land_use) >= 5)
rural_general_avgmon <- general_avgmon_full %>% filter(as.numeric(land_use) >= 5)
rural_native_monitors_avgmon <- rural_native_avgmon %>% filter(county_type == 1,
                                                           avg_mon > 0) 
rural_general_monitors_avgmon <- rural_general_avgmon %>% filter(county_type == 0,
                                                           avg_mon >0)
mean_mon_row_main <- data.frame(mean(rural_native_avgmon$avg_mon),
                                mean(rural_general_avgmon$avg_mon),
                                mean(rural_native_monitors_avgmon$avg_mon),
                                mean(rural_general_monitors_avgmon$avg_mon))
names(mean_mon_row_main) <- names(num_counties_row_main)

sd_mon_row_main <- data.frame(sd(rural_native_avgmon$avg_mon),
                                sd(rural_general_avgmon$avg_mon),
                                sd(rural_native_monitors_avgmon$avg_mon),
                                sd(rural_general_monitors_avgmon$avg_mon))
names(sd_mon_row_main) <- names(num_counties_row_main)

# Rural counties: monitor mean concentrations & SD
rural_native_avgexp <- native_avg_exp %>% filter(as.numeric(land_use) >= 5)
rural_general_avgexp <- general_avg_exp %>% filter(as.numeric(land_use) >= 5)

# new variables for PM2.5 concentrations in counties with monitors
monitors_rural_avgexp <- rural_native_avgexp
monitors_general_avgexp <- rural_general_avgexp

avg_PM25_row_main <-data.frame(mean(rural_native_avgexp$avg_pm_all),
                               mean(rural_general_avgexp$avg_pm_all),
                               mean(monitors_rural_avgexp$avg_pm_all),
                               mean(monitors_general_avgexp$avg_pm_all))
names(avg_PM25_row_main) <- names(num_counties_row_main) 

sd_PM25_row_main <- data.frame(sd(rural_native_avgexp$avg_pm_all),
                               sd(rural_general_avgexp$avg_pm_all),
                               sd(monitors_rural_avgexp$avg_pm_all),
                               sd(monitors_general_avgexp$avg_pm_all))
names(sd_PM25_row_main) <- names(num_counties_row_main) 
# Rural counties: model mean concentrations & SD
rural_native_modelexp <- avg_modelPM25 %>% filter(county_type == 1,
                                                  as.numeric(land_use) >= 5)
rural_general_modelexp <- avg_modelPM25 %>% filter(county_type == 0,
                                                   as.numeric(land_use) >= 5)
rural_native_monitor_modelexp <- avg_modelPM25 %>% inner_join(native_avg_exp) %>%
  filter(county_type == 1, as.numeric(land_use) >= 5)
rural_general_monitor_modelexp <- avg_modelPM25 %>% inner_join(monitors_general_avgexp) %>%
  filter(county_type == 0, as.numeric(land_use) >= 5)


avg_modelPM25_row_main <-data.frame(mean(rural_native_modelexp$avg_modelpm),
                               mean(rural_general_modelexp$avg_modelpm),
                               mean(rural_native_monitor_modelexp$avg_modelpm),
                               mean(rural_general_monitor_modelexp$avg_modelpm))
names(avg_modelPM25_row_main) <- names(num_counties_row_main) 

sd_modelPM25_row_main <- data.frame(sd(rural_native_modelexp$avg_modelpm),
                               sd(rural_general_modelexp$avg_modelpm),
                               sd(rural_native_monitor_modelexp$avg_modelpm),
                               sd(rural_general_monitor_modelexp$avg_modelpm))
names(sd_modelPM25_row_main) <- names(num_counties_row_main) 

#covariate mean rows! &SD
  # population density
popdensity_row_main <- data.frame(mean(rural_native_avgmon$pop_density),
                                  mean(rural_general_avgmon$pop_density),
                                  mean(rural_native_monitors_avgmon$pop_density),
                                  mean(rural_general_monitors_avgmon$pop_density))
names(popdensity_row_main) <- names(num_counties_row_main) 

sd_popden_main <- data.frame(sd(rural_native_avgmon$pop_density),
                                  sd(rural_general_avgmon$pop_density),
                                  sd(rural_native_monitors_avgmon$pop_density),
                                  sd(rural_general_monitors_avgmon$pop_density))
names(sd_popden_main) <- names(num_counties_row_main) 

  # hh income
hhinc_row_main <- data.frame(mean(rural_native_avgmon$hh_income),
                             mean(rural_general_avgmon$hh_income),
                             mean(rural_native_monitors_avgmon$hh_income),
                             mean(rural_general_monitors_avgmon$hh_income))
names(hhinc_row_main) <- names(num_counties_row_main) 
sd_hhinc_main <- data.frame(sd(rural_native_avgmon$hh_income),
                             sd(rural_general_avgmon$hh_income),
                             sd(rural_native_monitors_avgmon$hh_income),
                             sd(rural_general_monitors_avgmon$hh_income))
names(sd_hhinc_main) <- names(num_counties_row_main) 

  # ADI
adi_row_rural_25 <- data.frame(quantile(rural_native_avgmon$mean_nat_adi, 0.25, na.rm = T),
                                  quantile(rural_general_avgmon$mean_nat_adi, 0.25, na.rm = T),
                                  quantile(rural_native_monitors_avgmon$mean_nat_adi, 0.25, na.rm = T),
                                  quantile(rural_general_monitors_avgmon$mean_nat_adi, 0.25, na.rm = T))
names(adi_row_rural_25) <- names(num_counties_row_main) 

adi_rural_75 <- data.frame(quantile(rural_native_avgmon$mean_nat_adi, 0.75, na.rm = T),
                                  quantile(rural_general_avgmon$mean_nat_adi, 0.75, na.rm = T),
                                  quantile(rural_native_monitors_avgmon$mean_nat_adi, 0.75, na.rm = T),
                                  quantile(rural_general_monitors_avgmon$mean_nat_adi, 0.75, na.rm = T))
names(adi_rural_75) <- names(num_counties_row_main) 

#round mean rows first
num_counties_row_main[] <- lapply(num_counties_row_main,as.integer)
mean_mon_row_main <- format(round(mean_mon_row_main, 1), nsmall = 1)
avg_PM25_row_main <- format(round(avg_PM25_row_main, 1), nsmall = 1)
avg_modelPM25_row_main <-format(round(avg_modelPM25_row_main, 1), nsmall = 1)
popdensity_row_main[] <- lapply(popdensity_row_main,as.integer)
hhinc_row_main[] <- lapply(hhinc_row_main,as.integer)
adi_row_rural_25[] <- format(round(adi_row_rural_25, 1), nsmall = 1)

#round sd rows
sd_mon_row_main <- format(round(sd_mon_row_main, 1), nsmall = 1)
sd_PM25_row_main <- format(round(sd_PM25_row_main, 1), nsmall = 1)
sd_modelPM25_row_main <-format(round(sd_modelPM25_row_main, 1), nsmall = 1)
sd_popden_main[] <- lapply(sd_popden_main, as.integer)
sd_hhinc_main[] <- lapply(sd_hhinc_main, as.integer)
adi_rural_75[] <- format(round(adi_rural_75, 1), nsmall = 1)


#make list of rows
rows_list_main_rural <- list(num_counties_row_main,
      mean_mon_row_main,
      avg_PM25_row_main,
      avg_modelPM25_row_main,
      popdensity_row_main,
      hhinc_row_main,
      adi_row_rural_25)
rows_list_main_rural
#rbind all rows together
table_1_main_rural <- do.call(rbind, rows_list_main_rural)
table_1_main_rural
row.names(table_1_main_rural) <- c("Number of Counties",
                             "Average Monitor Count per year",
                        "Mean Monitor PM2.5 Concentration",
                        "Mean Model PM2.5 Concentrations",
                        "Population Density",
                        "HH Income",
                        "ADI 25th Percentile")
table_1_main_rural
# table1_rural_dt <- setDT(table_1_main_rural)
# main_table1_rural <- xtable(table1_rural_dt)
# main_table1_rural

#sd table (to manually input to table 1)
row_list_sd_rural <- list(sd_mon_row_main,
                    sd_PM25_row_main,
                    sd_modelPM25_row_main,
                    sd_popden_main,
                    sd_hhinc_main,
                    adi_rural_75)
table_1_sd_rural <- do.call(rbind,row_list_sd_rural)
row.names(table_1_sd_rural) <- c("SD Mean Monitor Count",
                           "SD Monitor PM2.5",
                        "SD Model PM2.5",
                        "Population Density",
                        "HH Income",
                        "ADI 75th Percentile")
table_1_sd_rural
# table1_sd_rural <- setDT(table_1_sd_rural)
# main_table_sd_rural <- xtable(table1_sd_rural)
# main_table_sd_rural
```

### 7/28/21 REVISION: To add row for mean % AI/AN by county type

```{r ALL COUNTIES}
native_census
prop_ai_county = all_native_counties %>% 
  rbind(all_general_counties) %>% 
  left_join(native_census) 

# Avg % AI for Non-AI counties
prop_ai_county %>% 
  filter(county_type == 0) %>% 
  pull(native_prop) %>% 
  mean()

# SD % AI for Non-AI counties
prop_ai_county %>% 
  filter(county_type == 0) %>% 
  pull(native_prop) %>% 
  sd()

# Avg % AI for AI counties
prop_ai_county %>% 
  filter(county_type == 1) %>% 
  pull(native_prop) %>% 
  mean()

# SD % AI for AI counties
prop_ai_county %>% 
  filter(county_type == 1) %>% 
  pull(native_prop) %>% 
  sd()

# For counties with monitors
# Avg % AI for AI counties
native_avg_exp %>% 
  pull(native_prop) %>% 
  mean()

#SD % AI for AI counties
native_avg_exp %>% 
  pull(native_prop) %>% 
  sd() 

# Avg % AI for non-AI counties
general_avg_exp  %>% 
  pull(native_prop) %>% 
  mean()
#SD % AI for non-AI counties
general_avg_exp  %>% 
  pull(native_prop) %>% 
  sd() 
```

```{r RURAL COUNTIES}
# all rural counties
rural_native %>% 
  pull(native_prop) %>% 
  mean()

rural_native %>% 
  pull(native_prop) %>% 
  sd()

rural_general %>% 
  pull(native_prop) %>% 
  mean()

rural_general %>% 
  pull(native_prop) %>% 
  sd()

# only counties with monitors
rural_native_avgexp %>% 
  pull(native_prop) %>% 
  mean()

rural_native_avgexp %>% 
  pull(native_prop) %>% 
  sd()

rural_general_avgexp %>% 
  pull(native_prop) %>% 
  mean()

rural_general_avgexp %>% 
  pull(native_prop) %>% 
  sd()
```
