---
title: 'Part 6: Descriptive Map'
author: "Maggie Li (ml4424)"
date: "10/12/2020"
output: html_document
---

```{r}
library(tidyverse)
library(leaflet)
library(htmltools)
```

## Descriptive Map of all the AI Counties

### Define AI counties as an sp file, by joining county lists of all 3 classifications with a county shapefile
```{r}
#map all native counties; to get sense of location to compare to monitor coverage
counties <- readOGR("Data/cb_2018_us_county_500k/cb_2018_us_county_500k.shp")
summary(counties)

#census
census_counties <- census %>%
  dplyr::rename(GEOID = County)
census_counties$State <- substr(census_counties$GEOID, 0, 2)
census_counties$State <- as.numeric(census_counties$State)
census_counties <- census_counties %>% filter(State <= 56,
                                              State != 2,
                                              State != 15)
census_counties
census_shp <- merge(counties, census_counties, by = 'GEOID', all = FALSE, duplicateGeoms = TRUE)
length(census_counties$GEOID)
length(census_shp$GEOID) #excludes 1 county

#cluster
cluster_counties <- cluster %>%
  dplyr::rename(GEOID = County)
cluster_counties$State <- substr(cluster_counties$GEOID, 0, 2)
cluster_counties$State <- as.numeric(as.character(cluster_counties$State))
cluster_counties <- cluster_counties %>% filter(State <= 56,
                                              State != 2,
                                              State != 15)
cluster_counties
cluster_shp <- merge(counties, cluster_counties, by = 'GEOID', 
                     all = FALSE, duplicateGeoms = TRUE)
summary(cluster_shp)
length(cluster_counties$GEOID)
length(cluster_shp$GEOID) 

#reservation
reservation_counties <- reservation %>% dplyr::rename(GEOID = County)
reservation_counties$State <- substr(reservation_counties$GEOID, 0, 2)
reservation_counties$State <- as.numeric(reservation_counties$State)
reservation_counties <- reservation_counties %>% 
  filter(State <= 56,
         State != 2,
         State != 15) 
unique(reservation_counties$GEOID)
reservation_shp <- merge(counties, reservation_counties, by = 'GEOID', 
                 all = FALSE, duplicateGeoms = TRUE)
length(reservation_counties$GEOID)
length(reservation_shp$GEOID) 
```

### Define monitor locations on AI counties to include in map

I differentiated between monitors active for 1-10 years and 11-19 years.
**July 30 update:** Plotted all PM2.5 monitors as just black points (rather than differentiating between monitors active for less than half or half or greater study years) so this section is commented out

```{r}
all_native_counties #recall df of distinct AI-classified counties

active_native_mon <- list(length(PM25_all))
for (i in 1:length(PM25_all)){
  active_native_mon[[i]] <- inner_join(PM25_all[[i]], all_native_counties)
}
agg_native_mon <- do.call(rbind, active_native_mon)
agg_native_mon # df of any active monitors across all study years
# 
# #filter by less than half or more than half of study years active
# agg_native_mon <- agg_native_mon %>% 
#   dplyr::select(-annual_mean)
# agg_native_mon <- agg_native_mon %>% dplyr::count(County, Latitude, Longitude) #new column n counts how many years each monitor (denoted by lat lon) is active
# agg_native_mon <- agg_native_mon %>% mutate(active = case_when(
#   n <= 10 ~ "1",
#   n > 10 ~ "2"
# )) #new column 'active' denotes monitors active for 1-10 years (active = 1) and those that are active for 11-19 years (active = 2)
# agg_native_mon
# agg_native_mon$State <- substr(agg_native_mon$County, 0, 2)
# #need to recode State as numeric, for random effects
# agg_native_mon$State <- as.numeric(agg_native_mon$State)
# #filter out territories, AL and HI
# agg_native_mon <- agg_native_mon %>% filter(State <= 56,
#                                       State != 2,
#                                       State != 15)
# view(agg_native_mon) #check that 'active' column should have entry 1 if 'n' column is less than or equal to 10, entry 2 if 'n' column greater than 10
```

### Map counties and monitor locations in Leaflet

Note: I'll probably end up mapping it in ggplot2 for the paper, so this is just a placeholder figure in the manuscript, just to make sure counties are showing up correctly.

8/5 Update: Map of native counties and ALL PM2.5 monitors
```{r}
# all_monitors <- do.call(rbind, PM25_all)
# all_unique_mon <- unique(all_monitors[c("Latitude", "Longitude")])
# # some fancy schmancy code to denote specific colors/shapes for map and legend symbols
# # different color polygons for the different county classifications and black points to symbolize locations of monitors
# colors <- c("#8dd3c7",  "#ffffb3", "#fb8072", "#000000")
# labels <- c("Native Census Population >5% County", 
#                       "Rural Wallace et al. Cluster",
#                       "County within Reservation",
#             "PM2.5 Monitor")
# sizes <- c(10, 10, 10, 5)
# shapes <- c("square", "square", "square", "circle")
# borders <- c("#8dd3c7",  "#ffffb3", "#fb8072", "#000000")
# addLegendCustom <- function(map, colors, labels, sizes, shapes, borders, opacity = 0.8, title){
#           make_shapes <- function(colors, sizes, borders, shapes) {
#               shapes <- gsub("circle", "50%", shapes)
#               shapes <- gsub("square", "0%", shapes)
#               paste0(colors, "; width:", sizes, "px; height:", sizes, 
#                      "px; border:3px solid ", borders, "; border-radius:", shapes)
#           }
#           make_labels <- function(sizes, labels) {
#               paste0("<div style='display: inline-block;height: ", 
#                      sizes, "px;margin-top: 4px;line-height: ", 
#                      sizes, "px;'>", labels, "</div>")
#           }
# 
#           legend_colors <- make_shapes(colors, sizes, borders, shapes)
#           legend_labels <- make_labels(sizes, labels)
# 
#           return(addLegend(map, colors = legend_colors, labels = legend_labels, opacity = opacity, title = title,
#                            position = "topright"))}
# native_allmon_map <- leaflet() %>%
#   addProviderTiles("CartoDB.PositronNoLabels") %>%
#   addPolygons(data = census_shp, color = "#8dd3c7",
#               fillOpacity = 0.8, weight = 3) %>%
#   addPolygons(data = cluster_shp, color = "#ffffb3", 
#               fillOpacity = 0.8, weight = 3) %>%
#   addPolygons(data = reservation_shp, color = "#fb8072",
#               fillOpacity = 0.8, weight = 3) %>%
#   addLegendCustom(colors, labels, sizes, shapes, borders,
#                   title = "American Indian Populated Counties") %>%
#   addCircleMarkers(data = all_unique_mon,
#                    lng = ~Longitude,
#                    lat = ~Latitude,
#                    radius = 1.5,
#                    stroke = FALSE, fillOpacity = 0.85,
#                    color = "black")
# native_allmon_map
```


Map of native counties and only PM2.5 monitors in native counties
```{r}
# # some fancy schmancy code to denote specific colors/shapes for map and legend symbols
# # different color polygons for the different county classifications and black points to symbolize locations of monitors
# colors <- c("#8dd3c7",  "#ffffb3", "#fb8072", "#000000")
# labels <- c("Native Census Population >5% County", 
#                       "Rural Wallace et al. Cluster",
#                       "County within Reservation",
#             "PM2.5 Monitor")
# sizes <- c(10, 10, 10, 5)
# shapes <- c("square", "square", "square", "circle")
# borders <- c("#8dd3c7",  "#ffffb3", "#fb8072", "#000000")
# addLegendCustom <- function(map, colors, labels, sizes, shapes, borders, opacity = 0.8, title){
#           make_shapes <- function(colors, sizes, borders, shapes) {
#               shapes <- gsub("circle", "50%", shapes)
#               shapes <- gsub("square", "0%", shapes)
#               paste0(colors, "; width:", sizes, "px; height:", sizes, 
#                      "px; border:3px solid ", borders, "; border-radius:", shapes)
#           }
#           make_labels <- function(sizes, labels) {
#               paste0("<div style='display: inline-block;height: ", 
#                      sizes, "px;margin-top: 4px;line-height: ", 
#                      sizes, "px;'>", labels, "</div>")
#           }
# 
#           legend_colors <- make_shapes(colors, sizes, borders, shapes)
#           legend_labels <- make_labels(sizes, labels)
# 
#           return(addLegend(map, colors = legend_colors, labels = legend_labels, opacity = opacity, title = title,
#                            position = "topright"))}
# native_counties_map <- leaflet() %>%
#   addProviderTiles("CartoDB.PositronNoLabels") %>%
#   addPolygons(data = census_shp, color = "#8dd3c7",
#               fillOpacity = 0.8, weight = 3) %>%
#   addPolygons(data = cluster_shp, color = "#ffffb3", 
#               fillOpacity = 0.8, weight = 3) %>%
#   addPolygons(data = reservation_shp, color = "#fb8072",
#               fillOpacity = 0.8, weight = 3) %>%
#   addLegendCustom(colors, labels, sizes, shapes, borders,
#                   title = "American Indian Populated Counties") %>%
#   addCircleMarkers(data = agg_native_mon,
#                    lng = ~Longitude,
#                    lat = ~Latitude,
#                    radius = 1.5,
#                    stroke = FALSE, fillOpacity = 0.85,
#                    color = "black")
# native_counties_map
```

Old code for differentiating between monitors active for less than half or greater than or equal than half of study years below:
```{r}
# # make palette for monitor colors based on years active status
# activeyrs_pal <- colorFactor(c("#80b1d3", "#fdb462"), domain = c("1", "2"))
# pal_native_mon <- colorNumeric(
#   palette = "inferno", 
#   domain = agg_native_mon$n,
#   reverse = TRUE)
# 
# # some fancy schmancy code to denote specific colors/shapes for map and legend symbols
# colors <- c("#8dd3c7",  "#ffffb3", "#fb8072", "#80b1d3", "#fdb462")
# labels <- c("Native Census Population >5% County", 
#                       "Rural Wallace et al. Cluster",
#                       "County within Reservation",
#             "Monitors Active Less than or Equal to Half of Study Years",
#               "Monitors Active Greater than Half of Years")
# sizes <- c(10, 10, 10, 5, 5)
# shapes <- c("square", "square", "square", "circle", "circle")
# borders <- c("#8dd3c7",  "#ffffb3", "#fb8072", "#80b1d3", "#fdb462")
# addLegendCustom <- function(map, colors, labels, sizes, shapes, borders, opacity = 0.8, title){
#           make_shapes <- function(colors, sizes, borders, shapes) {
#               shapes <- gsub("circle", "50%", shapes)
#               shapes <- gsub("square", "0%", shapes)
#               paste0(colors, "; width:", sizes, "px; height:", sizes, 
#                      "px; border:3px solid ", borders, "; border-radius:", shapes)
#           }
#           make_labels <- function(sizes, labels) {
#               paste0("<div style='display: inline-block;height: ", 
#                      sizes, "px;margin-top: 4px;line-height: ", 
#                      sizes, "px;'>", labels, "</div>")
#           }
# 
#           legend_colors <- make_shapes(colors, sizes, borders, shapes)
#           legend_labels <- make_labels(sizes, labels)
# 
#           return(addLegend(map, colors = legend_colors, labels = legend_labels, opacity = opacity, title = title,
#                            position = "bottomleft"))}
# native_counties_map <- leaflet() %>%
#   addProviderTiles(providers$CartoDB.Positron) %>%
#   addPolygons(data = census_shp, color = "#8dd3c7",
#               fillOpacity = 0.8, weight = 3) %>%
#   addPolygons(data = cluster_shp, color = "#ffffb3", 
#               fillOpacity = 0.8, weight = 3) %>%
#   addPolygons(data = reservation_shp, color = "#fb8072",
#               fillOpacity = 0.8, weight = 3) %>%
#   addLegendCustom(colors, labels, sizes, shapes, borders,
#                   title = "American Indian Populated Counties") %>%
#   addCircleMarkers(data = agg_native_mon,
#                    lng = ~Longitude,
#                    lat = ~Latitude,
#                    radius = 3,
#                    stroke = FALSE, fillOpacity = 0.85,
#                    color = ~activeyrs_pal(agg_native_mon$active))
# native_counties_map
```
### Map with just counties locations (no monitors)

```{r}
# some fancy schmancy code to denote specific colors/shapes for map and legend symbols
# different color polygons for the different county classifications and black points to symbolize locations of monitors
colors <- c("#8dd3c7",  "#ffffb3", "#fb8072")
labels <- c("Native Census Population >5% County", 
                      "Rural Wallace et al. Cluster",
                      "County within Reservation")
sizes <- c(10, 10, 10)
shapes <- c("square", "square", "square")
borders <- c("#8dd3c7",  "#ffffb3", "#fb8072")
addLegendCustom <- function(map, colors, labels, sizes, shapes, borders, opacity = 0.8, title){
          make_shapes <- function(colors, sizes, borders, shapes) {
              shapes <- gsub("circle", "50%", shapes)
              shapes <- gsub("square", "0%", shapes)
              paste0(colors, "; width:", sizes, "px; height:", sizes, 
                     "px; border:3px solid ", borders, "; border-radius:", shapes)
          }
          make_labels <- function(sizes, labels) {
              paste0("<div style='display: inline-block;height: ", 
                     sizes, "px;margin-top: 4px;line-height: ", 
                     sizes, "px;'>", labels, "</div>")
          }

          legend_colors <- make_shapes(colors, sizes, borders, shapes)
          legend_labels <- make_labels(sizes, labels)

          return(addLegend(map, colors = legend_colors, 
                           labels = legend_labels, 
                           opacity = opacity, 
                           title = title,
                           position = "topright"))}
no_monitors_map <- leaflet() %>%
  addProviderTiles("CartoDB.PositronNoLabels") %>%
  addPolygons(data = census_shp, color = "#8dd3c7",
              fillOpacity = 0.8, weight = 3) %>%
  addPolygons(data = cluster_shp, color = "#ffffb3", 
              fillOpacity = 0.8, weight = 3) %>%
  addPolygons(data = reservation_shp, color = "#fb8072",
              fillOpacity = 0.8, weight = 3) %>%
  addLegendCustom(colors, labels, sizes, shapes, borders,
                  title = "American Indian Populated Counties")

##Adjust size of legend
browsable(
 tagList(
    list(
       tags$head(
          tags$style(
             ".leaflet .legend {
               line-height: 15px;
               font-size: 15px;
               }",
            ".leaflet .legend i{
              width: 15px;
              height: 15px;
               }"
          )
       ),
     no_monitors_map)))

```

### Show overlap of AI county definitions (data visualized in venn diagram in paper to double check with)
```{r}
# Non-restricted for each definition (main analysis)
census_counties <- native_census_counties %>% 
  dplyr::select(County) 
cluster_counties <- native_rural_counties %>% 
  dplyr::select(County) 
reservation_counties <- reservation_counties %>%
  dplyr::select(County)

#three-way overlap
all_overlap <- cluster_counties %>% 
  inner_join(census_counties) %>% 
  inner_join(reservation_counties)
nrow(all_overlap) #number of counties that fit all three definitions

#two-way overlap
cluster_census <- cluster_counties %>% 
  inner_join(census_counties) %>% anti_join(all_overlap)
nrow(cluster_census)

cluster_reservation <- cluster_counties %>%
  inner_join(reservation_counties) %>% anti_join(all_overlap)
nrow(cluster_reservation)

census_reservation <- census_counties %>%
  inner_join(reservation_counties) %>% anti_join(all_overlap)
nrow(census_reservation)

#exclusive to each definition
census_only <- census_counties %>%
  anti_join(census_reservation) %>%
  anti_join(cluster_census) %>%
  anti_join(all_overlap)
nrow(census_only)

cluster_only <- cluster_counties %>%
  anti_join(cluster_census) %>%
  anti_join(cluster_reservation) %>%
  anti_join(all_overlap)
nrow(cluster_only)

reservation_only <- reservation_counties %>%
  anti_join(cluster_reservation) %>%
  anti_join(census_reservation) %>%
  anti_join(all_overlap)
nrow(reservation_only)
```

