---
title: "Part 5: Statistical Analysis"
author: "Maggie Li (ml4424)"
date: "10/16/2020"
output: html_document
---

```{r}
library(dplyr)
library(lme4)
library(stargazer)
library(gridExtra)
library(grid)
library(ggplot2)
```

### 8/13/2021 update: corrected confidence intervals to 1.96*std_err in graphs

## Linear Mixed Effects Model for Outcome = Monitor and Model PM2.5 concentrations

### Define lmer function
I am defining functions to run the following linear mixed regressions on the data, testing main effects only:
-Model 1: Crude (AI/Non-AI and Year only)
-Model 2: Partial (further adjusted for population density)
-Model 3: Full (further adjusted for both pop density and HH income)

I also run Model 3 to test for an interaction effect between AI/Non-AI classification and year.
-Model 3X: Full model with interaction term for county_type*year


```{r}
#Define referent category for binary "exposure" variable county_type
# Main Analysis:
all_ctyear_exp$county_type <- relevel(all_ctyear_exp$county_type,
                                         ref = "0")

# Check linearity of year, hh_inc, pop_density thru penalized spline
# Model with spline
library(mgcv)
library(gamm4)

# gamm with nested random intercepts for counties within states
gamm.nested = gamm4(annual_mean_all ~ county_type +
                      s(I(scale(hh_income))) + 
                      s(I(as.numeric(popd_q))) +
                      s(I(as.numeric(as.character(year))-2000)), # AI county is exposure of interest and linear term, confounders modeled as splines to check for linearity
                    random = ~(1|State/County), #nested random intercept
                    data = all_ctyear_exp)
summary(gamm.nested$gam) # view effect estimate of linear exposure term, edf of confounders modeled by splines
plot(gamm.nested$gam, pages = 1)

# only random intercept by state
# all models without pop density have small standard errors
gamm.eg <- gamm(annual_mean_all ~ county_type +
                  s(I(scale(hh_income))) + s(I(as.numeric(popd_q))) +
                  s(I(as.numeric(as.character(year))-2000)),
                random = list(State = ~1),
                      data = all_ctyear_exp)

summary(gamm.eg$gam)
plot(gamm.eg$gam, pages=1) #population density and HH income are non-linear, so we adjust for them using deciles

# Define functions for lmer
# Crude model only adjusting for county_type and year
detrend_lmer_crude <- function(data){
  exp_lme <- lmer(annual_mean_all ~ county_type + year +
                    (1|State/County),
                  data = data)
  return(exp_lme)}

# Partial model, adjusted for popd
detrend_lmer_partial <- function(data){
  exp_lme <- lmer(annual_mean_all ~ county_type + year +
                    popd_q +
                    (1|State/County),
                  data = data)
  return(exp_lme)}
# Full model with year as categorical variable
detrend_lmer_full <- function(data){
  exp_lme <- lmer(annual_mean_all ~ county_type + year +
                    popd_q + 
                    hhinc_q +
                    (1|State/County),
                  data = data, REML=FALSE)
  return(exp_lme)}

# Full model with year as categorical variable, interx between year and native 
detrend_lmer_full_catyr_intx <- function(data){
  exp_lme <- lmer(annual_mean_all ~ county_type + year +
                    popd_q + 
                    hhinc_q +
                    county_type*year +
                    (1|State/County),
                  data = data, REML=FALSE)
  return(exp_lme)}
```

### Monitor Data: Run regressions for main analysis (all counties) and sensitivity analysis (rural counties)

#### Main Analysis
```{r}
test = all_ctyear_exp %>% inner_join(climatezonefips)
test

# Full w/ InterX
summary(detrend_lmer_full_catyr_intx(test))

# Full w/ InterX with climate as variable
exp_lme_climate <- lmer(annual_mean_all ~ county_type + year +
                  popd_q + 
                  hhinc_q +
                  Climate_Zone +
                  county_type*year +
                  (1|State/County),
                data = test, REML=FALSE)
summary(exp_lme_climate)

# Crude
summary(detrend_lmer_crude(all_ctyear_exp))$coefficients

# Partial
summary(detrend_lmer_partial(all_ctyear_exp))$coefficients
# Full
summary(detrend_lmer_full(all_ctyear_exp))$coefficients
# Full w/ InterX
summary(detrend_lmer_full_catyr_intx(all_ctyear_exp))

# Model Output Table with three columns for crude, , full
stargazer(detrend_lmer_crude(all_ctyear_exp),
          detrend_lmer_partial(all_ctyear_exp),
          detrend_lmer_full(all_ctyear_exp),
          ci=T)

# # 8/25/20: read out all_ctyear_exp to send to Marianthi
# setwd("Data/intermediate_data")
# write_csv(all_ctyear_exp, "monitor_PM25_all.csv")
```

#### Rural Counties; Sensitivity Analysis


```{r}
# subset rural counties
rural_monitors <- all_ctyear_exp %>% filter(as.numeric(as.factor(land_use)) >= 5)

#Run on Monitor Data for rural counties
summary(detrend_lmer_full(rural_monitors))
summary(detrend_lmer_partial(rural_monitors))
summary(detrend_lmer_crude(rural_monitors))

#output latex table
stargazer(detrend_lmer_crude(rural_monitors),ci=T)
```

#### Plot time-varying effect of AI/Non-AI county type binary variable (county_type interaction effect with year)


Creating Variance-Covariance Matrices, Plotting model PM2.5 declines over time

I first create a variance-covariance matrix to pull out all of the variances for all of the study years after 2000:
var(x+y) = var(x) + var(y) + 2*covar(x,y) where x = main effect of native county type on PM2.5 concnetrations in 2000 (referent), y = main effect of native county type due to interaction with time in study years 2001-2018

Then I create a new data frame, pm_decline_monitor_all, which saves all of the total effect estimates (main effect + interaction effect) and their 95% confidence intervals. I use a loop to iterate through all the study years to fill in this table. Then I can graph this!

##### All Main Analysis Counties

```{r}
# create vcov matrix of main effects and interX

# 8/17/20 edit: first define variable for model function to avoid rerunning it a million times
main_interx_monitor <- detrend_lmer_full_catyr_intx(all_ctyear_exp)
summary(main_interx_monitor)
native_yr_vcov <- vcov(main_interx_monitor)[c(2,seq(39,56)), c(2,seq(39,56))]

# calculate all the variances for all the years i.e. var(x+y); should be 19 total entries
var_vector = c()
for (i in 2:19){
  var_vector[1] <- native_yr_vcov[1,1]
  var_vector[i] <- native_yr_vcov[1,1] + native_yr_vcov[i,i] + 2 * native_yr_vcov[i,1]
}
var_vector
sd_vector <- sqrt(var_vector)
length(sd_vector)
#matrix with 19 cols for 19 years, three rows: one for effect estimate of total effect per year (total effect = main effect + interx effect), one for CI lower, one for CI upper

## for 2000
pm_decline_monitor_all <- data.frame()
pm_decline_monitor_all[1,1] <- summary(main_interx_monitor)$coefficients[2,1]
pm_decline_monitor_all[1,2] <- summary(main_interx_monitor)$coefficients[2,1] - 1.96*sd_vector[1]
pm_decline_monitor_all[1,3] <- summary(main_interx_monitor)$coefficients[2,1] + 1.96*sd_vector[1]

# fill in matrix thru loop for every following year 2001-2018
yr_ct <- 39
for (i in 2:19){
  pm_decline_monitor_all[i,1] <- summary(main_interx_monitor)$coefficients[2,1]+
    summary(main_interx_monitor)$coefficients[yr_ct,1]
  
  pm_decline_monitor_all[i,2] <- summary(main_interx_monitor)$coefficients[2,1]+
    summary(main_interx_monitor)$coefficients[yr_ct,1] - 1.96*sd_vector[i]
  
  pm_decline_monitor_all[i,3] <- summary(main_interx_monitor)$coefficients[2,1]+
    summary(main_interx_monitor)$coefficients[yr_ct,1] + 1.96*sd_vector[i]
  yr_ct <- yr_ct + 1
}

pm_decline_monitor_all

colnames(pm_decline_monitor_all) <- c('estimate', 'cl_lower', 'cl_upper') # set col names

pm_decline_monitor_all$year <- seq(2000, 2018) # column for year


#PLOT OF TOTAL EFFECT OF NATIVE OVER TIME; updated 12/14 to be consistent with Figure 3 layout (slanted x-axis label)
monitor_interx_plot <-ggplot() + 
  theme_linedraw() + 
  geom_line(data = pm_decline_monitor_all,
            aes(x=year, y = estimate)) +
  geom_line(data=pm_decline_monitor_all,
            aes(x=year, y=cl_lower), linetype = "dashed") +
    geom_line(data=pm_decline_monitor_all,
            aes(x=year, y=cl_upper), linetype = "dashed") +
  labs(x = "Year",
       y = expression(paste("Mean Difference in ", PM[2.5], " (", mu, "g/", m^3, ")")),
       fill = "County Type") +
  ylim(-2.7, 1.7) +
  theme(plot.title = element_text(size = 18),
        axis.title = element_text(size = 16),
        axis.text = element_text(size = 12),
        # axis.title.x = element_blank(),
        axis.title.y = element_blank()
        ) +
  scale_x_continuous(breaks = seq(2000,2018,1), expand = c(0, 0)) + 
  guides(x =  guide_axis(angle = 45)) +
  geom_hline(yintercept=0, linetype="solid", color = "red") 
monitor_interx_plot 

# 12/14 for paper
grob_monitor <- grobTree(textGrob("B", x=0.05,  y=0.9, hjust=0,
  gp=gpar(fontsize=30)))

# Plot and save
monitor_interx_grob = monitor_interx_plot + annotation_custom(grob_monitor)
ggsave("figures/updated_12.7/monitor_interx.pdf", width = 6, height = 4, dpi = 300)


monitor_interx_plot
ggsave("presentations/layering_trends/monitor_interx.pdf", width = 6, height = 4, dpi = 300)
```

##### Rural Main Analysis Counties (Sensitivity Analysis)

```{r}
# 8/17/20 edit: first define variable for model function to avoid rerunning it a million times
rural_interx_monitor <- detrend_lmer_full_catyr_intx(rural_monitors)

# vcov matrix for all year combinations with native county type (one less p bc one less pop density top quantile category variable)
native_yr_vcov <- vcov(rural_interx_monitor)[c(2,seq(35,52)), c(2,seq(35,52))] # 4 less rows and columns bc missing one quantile of hinc, popdens from all study counties/main analysis 

# calculate all the variances for all the years i.e. var(x+y); should be 19 total entries
var_vector = c()
for (i in 2:19){
  var_vector[1] <- native_yr_vcov[1,1] # year 2000
  var_vector[i] <- native_yr_vcov[1,1] + native_yr_vcov[i,i] + 2 * native_yr_vcov[i,1] # every following year with additional interaction beta
}
var_vector
sd_vector <- sqrt(var_vector)
length(sd_vector)
#matrix with 19 cols for 19 years, three rows: one for effect estimate of total effect per year (total effect = main effect + interx effect), one for CI lower, one for CI upper
pm_decline_monitor_rural <- data.frame()

# for year 2000
pm_decline_monitor_rural[1,1] <- summary(rural_interx_monitor)$coefficients[2,1]
pm_decline_monitor_rural[1,2] <- summary(rural_interx_monitor)$coefficients[2,1] - 1.96*sd_vector[1]
pm_decline_monitor_rural[1,3] <- summary(rural_interx_monitor)$coefficients[2,1] + 1.96*sd_vector[1]

# fill in matrix thru loop for every following year 2001-2017
yr_ct <- 35
for (i in 2:19){
  pm_decline_monitor_rural[i,1] <- summary(rural_interx_monitor)$coefficients[2,1]+
    summary(rural_interx_monitor)$coefficients[yr_ct,1]
  pm_decline_monitor_rural[i,2] <- summary(rural_interx_monitor)$coefficients[2,1]+
    summary(rural_interx_monitor)$coefficients[yr_ct,1] - 1.96*sd_vector[i]
  pm_decline_monitor_rural[i,3] <- summary(rural_interx_monitor)$coefficients[2,1]+
    summary(rural_interx_monitor)$coefficients[yr_ct,1] + 1.96*sd_vector[i]
  yr_ct <- yr_ct + 1
}
pm_decline_monitor_rural
# set col names
colnames(pm_decline_monitor_rural) <- c('estimate', 'cl_lower', 'cl_upper')
pm_decline_monitor_rural$year <- seq(2000, 2018)

#PLOT OF TOTAL EFFECT OF NATIVE OVER TIME
monitor_interx_rural <-ggplot() + 
  theme_linedraw() + 
  geom_line(data = pm_decline_monitor_rural,
            aes(x=year, y = estimate)) +
  geom_line(data=pm_decline_monitor_rural,
            aes(x=year, y=cl_lower), linetype = "dashed") +
    geom_line(data=pm_decline_monitor_rural,
            aes(x=year, y=cl_upper), linetype = "dashed") +
  labs(x = "Year",
       y = expression(paste("Mean Difference in ", PM[2.5], " (", mu, "g/", m^3, ")")),
       fill = "County Type") +
  ylim(-2.7, 1.7) +
  theme(plot.title = element_text(size = 18),
        axis.title = element_text(size = 16),
        axis.text = element_text(size = 12),
        # axis.title.x = element_blank(),
        # axis.title.y = element_blank()
        ) +
  scale_x_continuous(breaks = seq(2000,2018,1), expand = c(0, 0)) + 
  guides(x =  guide_axis(angle = 45)) +
  geom_hline(yintercept=0, linetype="solid", color = "red") 
monitor_interx_rural 

grob <- grobTree(textGrob("B", x=0.05,  y=0.9, hjust=0,
  gp=gpar(fontsize=30)))

# Plot and save
monitor_interx_rural + annotation_custom(grob)
ggsave("figures/updated_12.7/monitor_rural_interx.png", width = 6, height = 4, dpi = 300)
```

### Model Data: Run regressions for main analysis (all counties) and sensitivity analysis (rural counties)


#### Clean Modeled PM25 data to run models in (7_stat_analysis, non-NE in 8_supplement)
```{r}
# Split population density and hhincome into deciles for model
model_PM25$popd_q <- cut(model_PM25$pop_density, quantile(model_PM25$pop_density, seq(0,1,0.1)), include.lowest = TRUE)
model_PM25$hhinc_q <- cut(model_PM25$hh_income, quantile(model_PM25$hh_income, seq(0,1,0.1)), include.lowest = TRUE)
sum(table(model_PM25$popd_q, exclude = NULL)) == dim(model_PM25)[1]

# edit cols to fit model function for linear mixed effects model
model_PM25$county_type <- as.factor(as.character(model_PM25$county_type))

  # set referent
model_PM25$county_type <- relevel(model_PM25$county_type,
                                         ref = "0")
model_PM25$Year <- as.factor(model_PM25$Year)
model_PM25 <- dplyr::rename(model_PM25, year=Year)

# rename column for PM25 to match predefined function arguments
model_lmer <-  dplyr::rename(model_PM25,annual_mean_all=PM25)
model_lmer
```

#### Main Analysis
```{r}
# Run LMER (crude, partial, adjusted)
summary(detrend_lmer_crude(model_lmer))$coefficients
summary(detrend_lmer_partial(model_lmer))$coefficients
summary(detrend_lmer_full(model_lmer))$coefficients

# All Models output table
stargazer(detrend_lmer_crude(model_lmer),
          detrend_lmer_partial(model_lmer),
          detrend_lmer_full(model_lmer),
          ci=T)

test = lmer(annual_mean_all ~ county_type*year +
                    (1|State/County),
                  data = model_lmer)
summary(test)
summary(detrend_lmer_full_catyr_intx(model_lmer))

## Distribution of variables
ai_pop_d = model_lmer %>% 
  filter(county_type == 1) %>% 
  pull(popd_q) 
non_ai_pop_d = model_lmer %>% 
  filter(county_type == 0) %>% 
  pull(popd_q) 
hist(as.numeric(non_ai_pop_d))
hist(as.numeric(ai_pop_d))

ai_hh_inc = model_lmer %>% 
  filter(county_type == 1) %>% 
  pull(hhinc_q) 
non_ai_hh_inc = model_lmer %>% 
  filter(county_type == 0) %>% 
  pull(hhinc_q) 
hist(as.numeric(non_ai_hh_inc))
hist(as.numeric(ai_hh_inc))

model_PM_climate = model_lmer %>% inner_join(climatezonefips)
model_PM_climate

# model w/o interX
modelpm_lme_climate <- lmer(annual_mean_all ~ county_type + year +
                  popd_q + 
                  hhinc_q +
                  Climate_Zone +
                  (1|State/County),
                data = model_PM_climate, REML=FALSE)
summary(modelpm_lme_climate)

# Full model w/ InterX
summary(detrend_lmer_full_catyr_intx(model_lmer))

# Full w/ InterX with climate as variable
modelpm_lme_climate_interx <- lmer(annual_mean_all ~ county_type + year +
                  popd_q + 
                  hhinc_q +
                  Climate_Zone +
                  county_type*year +
                  (1|State/County),
                data = model_PM_climate, REML=FALSE)
summary(modelpm_lme_climate_interx)

anova(detrend_lmer_full(model_lmer),
      detrend_lmer_full_catyr_intx(model_lmer))


```

#### Rural Counties; Sensitivity Analysis

```{r}
# subset rural counties
rural_model <- model_lmer %>% filter(as.numeric(as.factor(land_use)) >= 5)

#Run on Model Data for rural counties (main effects only model)
summary(detrend_lmer_full(rural_model))

#output latex table
stargazer(detrend_lmer_full(rural_model), ci=T)

unique(rural_model$popd_q)
unique(rural_model$hhinc_q)

# point estimate
summary(detrend_lmer_full(rural_model))$coefficients[2,1]
# 95% upper
summary(detrend_lmer_full(rural_model))$coefficients[2,1] + 1.96*summary(detrend_lmer_full(rural_model))$coefficients[2,2]

# 95% lower
summary(detrend_lmer_full(rural_model))$coefficients[2,1] - 1.96*summary(detrend_lmer_full(rural_model))$coefficients[2,2]
```

#### Plot time-varying effect of AI/Non-AI county type binary variable (county_type interaction effect with year)

##### All Main Analysis Counties
```{r}
# 8/17/20 edit: first define variable for model function to avoid rerunning it a million times
main_interx_modeled <- detrend_lmer_full_catyr_intx(model_lmer)
summary(main_interx_modeled)
# vcov matrix
native_yr_vcov <- vcov(main_interx_modeled)[c(2,seq(39,56)), c(2,seq(39,56))]
# calculate all the variances for all the years i.e. var(x+y); should be 19 total entries
var_vector = c()

for (i in 2:19){
  var_vector[1] <- native_yr_vcov[1,1]
  var_vector[i] <- native_yr_vcov[1,1] + native_yr_vcov[i,i] + 2 * native_yr_vcov[i,1]
}

sd_vector <- sqrt(var_vector)
sd_vector
#matrix with 19 cols for 19 years, three rows: one for effect estimate of total effect per year (total effect = main effect + interx effect), one for CI lower, one for CI upper
pm_decline_model_all <- data.frame()

# test: fill in values for 2001
pm_decline_model_all[1,1] <- summary(main_interx_modeled)$coefficients[2,1]
pm_decline_model_all[1,2] <- summary(main_interx_modeled)$coefficients[2,1] - 1.96*sd_vector[1]
pm_decline_model_all[1,3] <- summary(main_interx_modeled)$coefficients[2,1] + 1.96*sd_vector[1]

# fill in matrix thru loop for every following year 2002-2018
yr_ct <- 39
for (i in 2:19){
  pm_decline_model_all[i,1] <- summary(main_interx_modeled)$coefficients[2,1]+
    summary(main_interx_modeled)$coefficients[yr_ct,1]
  pm_decline_model_all[i,2] <- summary(main_interx_modeled)$coefficients[2,1]+
    summary(main_interx_modeled)$coefficients[yr_ct,1] - 1.96*sd_vector[i]
  pm_decline_model_all[i,3] <- summary(main_interx_modeled)$coefficients[2,1]+
    summary(main_interx_modeled)$coefficients[yr_ct,1] + 1.96*sd_vector[i]
  yr_ct <- yr_ct + 1
}
# set col names
colnames(pm_decline_model_all) <- c('estimate', 'cl_lower', 'cl_upper')
pm_decline_model_all$year <- seq(2000, 2018)
pm_decline_model_all

#PLOT OF TOTAL EFFECT OF NATIVE OVER TIME
model_interx_plot <-ggplot() + 
  theme_linedraw() + 
  geom_line(data = pm_decline_model_all,
            aes(x=year, y = estimate)) +
  geom_line(data=pm_decline_model_all,
            aes(x=year, y=cl_lower), linetype = "dashed") +
    geom_line(data=pm_decline_model_all,
            aes(x=year, y=cl_upper), linetype = "dashed") +
  # ggtitle(expression(paste("Modeled ", PM[2.5], " Difference in AI vs. Non-AI Populated Counties"))) +
  labs(x = "Year",
       y = expression(paste("Mean Difference in ", PM[2.5], " (", mu, "g/", m^3, ")")),
       fill = "County Type") +
  ylim(-2.7, 1.7) +
  theme(plot.title = element_text(size = 18),
        axis.title = element_text(size = 16),
        axis.text = element_text(size = 12),
        # axis.title.x = element_blank(),
        # axis.title.y = element_blank()
        ) +
  scale_x_continuous(breaks = seq(2000,2018,1), expand = c(0, 0)) + 
  guides(x =  guide_axis(angle = 45)) +
  geom_hline(yintercept=0, linetype="solid", color = "red") 
model_interx_plot 

# 12/14 for paper
grob <- grobTree(textGrob("A", x=0.05,  y=0.9, hjust=0,
  gp=gpar(fontsize=30)))

# Plot and save
model_interx_grob = model_interx_plot + annotation_custom(grob)
ggsave("figures/updated_12.7/model_interx.pdf", width = 6, height = 4, dpi = 300)

# save without grob
model_interx_plot
ggsave("presentations/layering_trends/model_interx.pdf", width = 6, height = 4)

library(ggpubr)
ggarrange(model_interx_grob, monitor_interx_grob)
ggsave("figures/updated_12.7/model_monitor_interx.pdf", width = 12, height = 4, dpi = 300)
# 
# # save without grob for F31 prelim data
# model_interx_plot
# ggsave("presentations/layering_trends/model_interx.png")
```
```{r}
# Generate a list of the aligned plot objects (measured and modeled pm2.5)
# library(cowplot)
# pl <- align_plots(monitor_interx_plot, model_interx_plot, align="v")
# p1
# ggdraw(pl[[1]])
# monitor_interx_plot
# model_interx_plot
```

##### Rural Main Analysis Counties only
```{r}
# 8/17/20 edit: first define variable for model function to avoid rerunning it a million times
rural_interx_modeled <- detrend_lmer_full_catyr_intx(rural_model)

# vcov matrix
native_yr_vcov <- vcov(rural_interx_modeled)[c(2,seq(39,56)), c(2,seq(39,56))]

# calculate all the variances for all the years i.e. var(x+y); should be 19 total entries
var_vector = c()
for (i in 2:19){
  var_vector[1] <- native_yr_vcov[1,1]
  var_vector[i] <- native_yr_vcov[1,1] + native_yr_vcov[i,i] + 2 * native_yr_vcov[i,1]
}
var_vector
sd_vector <- sqrt(var_vector)
length(sd_vector)
#matrix with 19 cols for 19 years, three rows: one for effect estimate of total effect per year (total effect = main effect + interx effect), one for CI lower, one for CI upper
pm_decline_model_rural <- data.frame()
pm_decline_model_rural[1,1] <- summary(rural_interx_modeled)$coefficients[2,1]
pm_decline_model_rural[1,2] <- summary(rural_interx_modeled)$coefficients[2,1] - 1.96*sd_vector[1]
pm_decline_model_rural[1,3] <- summary(rural_interx_modeled)$coefficients[2,1] + 1.96*sd_vector[1]

# fill in matrix thru loop for every following year
yr_ct <- 39
for (i in 2:19){
  pm_decline_model_rural[i,1] <- summary(rural_interx_modeled)$coefficients[2,1]+
    summary(rural_interx_modeled)$coefficients[yr_ct,1]
  
  pm_decline_model_rural[i,2] <- summary(rural_interx_modeled)$coefficients[2,1]+
    summary(rural_interx_modeled)$coefficients[yr_ct,1] - 1.96*sd_vector[i]
  
  pm_decline_model_rural[i,3] <- summary(rural_interx_modeled)$coefficients[2,1]+
    summary(rural_interx_modeled)$coefficients[yr_ct,1] + 1.96*sd_vector[i]
  yr_ct <- yr_ct + 1
}
pm_decline_model_rural
# set col names
colnames(pm_decline_model_rural) <- c('estimate', 'cl_lower', 'cl_upper')
pm_decline_model_rural$year <- seq(2000, 2018)

#PLOT OF TOTAL EFFECT OF NATIVE OVER TIME
model_interx_rural <-ggplot() + 
  theme_linedraw() + 
  geom_line(data = pm_decline_model_rural,
            aes(x=year, y = estimate)) +
  geom_line(data=pm_decline_model_rural,
            aes(x=year, y=cl_lower), linetype = "dashed") +
    geom_line(data=pm_decline_model_rural,
            aes(x=year, y=cl_upper), linetype = "dashed") +
  # ggtitle(expression(paste("Modeled ", PM[2.5], " Difference in AI vs. Non-AI Populated Counties"))) +
  labs(x = "Year",
       y = expression(paste("Mean Difference in ", PM[2.5], " (", mu, "g/", m^3, ")")),
       fill = "County Type") +
  ylim(-2.7, 1.7) +
  theme(plot.title = element_text(size = 18),
        axis.title = element_text(size = 16),
        axis.text = element_text(size = 12),
        # axis.title.x = element_blank(),
        # axis.title.y = element_blank()
        ) +
  scale_x_continuous(breaks = seq(2000,2018,1), expand = c(0, 0)) + 
  guides(x =  guide_axis(angle = 45)) +
  geom_hline(yintercept=0, linetype="solid", color = "red") 
model_interx_rural 

grob <- grobTree(textGrob("A", x=0.05,  y=0.9, hjust=0,
  gp=gpar(fontsize=30)))

# Plot and save
model_interx_rural + annotation_custom(grob)
ggsave("figures/updated_12.7/model_rural_interx.png", width = 6, height = 4, dpi = 300)
```


## Read out intermediate data files
```{r}
write_csv(all_ctyear_exp, "intermediate/monitor_ctyear.csv")
write_csv(rural_monitors, "intermediate/monitor_rural_ctyear.csv")
write_csv(model_lmer, "intermediate/model_ctyear.csv")
write_csv(rural_model, "intermediate/model_rural_ctyear.csv")
```
